### 2019-02-23 pge-mattis.tex
Add initial version

\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
\RequirePackage{fix-cm}
\documentclass[draft]{svjour3}
\smartqed
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[final]{graphicx}
\graphicspath{ {./images/} }
\usepackage{amsmath,amssymb}
\usepackage[final]{hyperref}
\usepackage{tikz}
\usepackage{environ}
\usetikzlibrary{shapes,arrows,positioning,calc}
\begin{document}

\title{Policy Gradient Estimation}
\subtitle{A survey on policy gradient estimators}

\author {Mattis Manfred K{\"a}mmerer \and Mahdi Enan \and Simon Kiefhaber}

\institute{Mattis Manfred K{\"a}mmerer \and Mahdi Enan \and Simon Kiefhaber \at TU Darmstadt}

\date{Received: date / Accepted: date}

\maketitle

\begin{abstract}
Policy gradient methods use a policy model to decide which actions maximize the expected accumulated reward received in a given environment. Policies can be deterministic but are often stochastic to accommodate for the nature of the given environment. One class of problems that is frequently chosen to evaluate these methods are continuous control problems. Depending on the environment the formal design of an optimal controller can be infeasible or very labor-intensive. Policy gradient methods can provide useful control models while aiming to optimize their data efficiency, thus requiring less resources for the learning process. The efficiency of these methods highly depends on the direction of change, i.e., the gradient they estimate. The goal of policy gradient estimation is to estimate a function that predicts how the policy should be changed to solve the reinforcement learning problem of maximizing the accumulated reward of an agent.  Thus, different functions for estimating the gradient have been proposed, some of which outperform others in specific areas. This survey gives a general introduction to policy gradient, and gives an overview of existing gradient estimation approaches.

\keywords{reinforcement learning \and policy gradients \and actor-critic \and advantage estimation \and continuous control }
\end{abstract}

\section{Introduction}
\label{intro}

Introductions to reinforcement learning often start by introducing value-based methods. They aim to gain the maximum accumulated reward in a Markov Decision Process (MDP). Value-based methods estimate the expected value of each action in each state, approximating a function that ranks actions in a given state, which is then used to derive a policy. Though this works well for problems with countable actions and states, its performance is not stable on continuous control problems. Even with optimal hyperparameters, the outcomes vary between separate trainings. \cite{DBLP:journals/corr/MetzIJD17} In constract to value-based methods, the other main class of model-free reinforcement learning algorithms is called policy-based methods, or policy gradient methods. Research has shown, that policy gradient methods can reach better performance in these environments. \cite{pmlr-v37-schulman15} Using a parameterized (often stochastic) policy to decide the next action, they can easily incorporate prior domain knowledge, and reuse some of the value-based approaches for improved performance, but require a lot of configuration to produce an effective agent for a specific environment. Also, they frequently require on-policy training and often converge slower in discrete settings. Options to model a policy gradient method are not only given by the policy model itself, but also in the way a policy improvement is computed. \cite{DBLP:journals/corr/SchulmanMLJA15} In this case, improvement means a step in the parameter space such that the policy under the new parameters will on average perform better than the old policy. Policy gradient estimation is the term we use to describe the process of computing the next step in parameter space. The goal is to estimate a function that predicts how the policy should be changed to maximize the accumulated reward of an agent in an MDP. Since this is the core problem of policy gradient methods, it is also the main topic of this paper.

In section \ref{sec:prel}, we give some preliminaries and describe the problem setup in detail. In section \ref{sec:pge}, we discuss different approaches to estimate the policy gradient. Using our insights from section \ref{sec:pge}, we derive the actor-critic framework in section \ref{sec:ac}, which harnesses some value-based approaches for improved gradient updates. Then, in section \ref{sec:natural}, we introduce some gradient ascent methods that build on the approaches given in \ref{sec:pge}, refining the gradient estimation by Fisher's information matrix to get the natural gradient. Finally, in section \ref{sec:outro}, we give a conclusion of the reviewed approaches, and discuss potential areas of improvement.

\section{Preliminaries}
\label{sec:prel}

\paragraph{Definitions.} We define states $s \in \mathbb{X}$, actions $a \in \mathbb{U}$, and rewards $r \in \mathbb{R}$. A trajectory $\rho_{\pi_\theta}^h := (s_0, a_0, $ $s_1, a_1, \dots, s_h, a_h)$ is generated by drawing $s_0 \sim p(x_0)$ according to the distribution over initial states $p(x_0)$, and successively sampling $a_t \sim \pi_\theta(a_t|s_t)$ according to the policy $\pi_\theta$, and $s_{t+1} \sim p(s_{t+1}|s_t,a_t)$ until the horizon $h$ is reached. At each time step, we receive a reward according to $r_t = r(s_t, a_t, s_{t+1})$. A trajectory can also be called \textit{roll-out} or \textit{episode}, though the term episode implies it ends in a terminal state.

We assume a Markov Decision Process (MDP), meaning the proability distribution of the next states is independent of past states $s_{0:t-1}$ given the present state $s_t$ and action $a_t$, 
\begin{equation}
	p(s_{t+1}|s_t,a_t)=p(s_{t+1}|s_{0:t},a_{0:t}).
\end{equation}
Where we define $i:j$ with $i,j \in \mathbb{N}, i < j$ as an index over all integers from $i$ to $j$, i.e., $s_{i:j} := s_i, s_{i+1}, \dots, s_j$.

Distributions often used as policies are Gibbs policies $\pi_\theta(a|s) = \frac{exp(\phi(s,a)^T\theta)}{\sum_b exp(\phi(s,b)^T\theta)}$ \cite{Sutton:1999:PGM:3009657.3009806,Bagnell2004LearningD} for discrete problems, and Gaussian policies $\pi_\theta(a|s) = N(\phi(s,a)^T\phi_1,\phi_2)$ for continuous problems, where $\phi_2$ is an exploration parameter \cite{Williams92simplestatistical,peter:article:1996}. $\phi(s,a)$ is the vector of basis functions on the state-action pair.

\paragraph{Policy gradient.} Our goal with respect to episodes is to maximize the expectation of the total reward $\mathbb{E}_{s\sim p(s|\theta)}\left[\mathcal{R}^\infty\right]$. The total reward in the horizon $h$ for the non-discounted problem is $\sum_{t=0}^{h} r_{t}$. For most cases, a discount factor $\gamma \in [0,1)$ is added to ensure that this sum is finite for the infinite horizon $h\to\infty$. Intuitively, this reflects the idea that the relevance of later actions declines, and reduces variance during learning. The discounted total reward is

\begin{equation}
  \mathcal{R}^h := \sum_{t=0}^{h} \gamma^t r_t.
  \label{eqn:acc-reward}
\end{equation}

Since we have only empirical knowledge of the performance of the policy, we need to approximate an optimal policy by estimating a gradient. Thus, we search $\nabla_\theta J(\theta) := \nabla_\theta \mathbb{E}_{s\sim p(s|\theta)}\left[\psi(s,a)\right]$, to make a policy gradient step according to $\theta_{k+1} = \theta_k + \alpha_k \nabla_\theta J(\theta)$, where $\psi$ denotes some score function, see section \ref{sec:pge}, and $\alpha_k$ denotes a learning rate. It is shown that for $\lim_{k\to\infty}\alpha_k = 0$, and $\sum_{k=0}^\infty \alpha_k$ we are guaranteed to converge to a local optimum \cite{Sutton:1999:PGM:3009657.3009806}.

\paragraph{Likelihood-ratio gradients.} Using the definition of the expectation, we derive
\begin{align*}
  \nabla_\theta \mathbb{E}_{s\sim p(s|\theta)}\left[\psi(s,a)\right] &= \nabla_\theta \int_\mathbb{X}{p(s|\theta)\int_\mathbb{U}{\pi_\theta(a|s)\psi(s,a) da}ds} \\
  &= \int_\mathbb{X}{p(s|\theta)\int_\mathbb{U}{\nabla_\theta {\pi_\theta(a|s)} \psi(s,a) da}ds} \\
  &= \int_\mathbb{X}{p(s|\theta)\int_\mathbb{U}{\pi_\theta(a|s)\nabla_\theta{\log\pi_\theta(a|s)} \psi(s,a) da}ds} .
\end{align*}
Where we used $\nabla_\theta \pi_\theta(a|s) = \pi_\theta(a|s)\nabla_\theta\log\pi_\theta(a|s)$, obtained from the likelihood ratio, which gives us the likelihood-ratio gradient,
\begin{equation}
  \nabla_\theta \mathbb{E}_{\pi_\theta}\left[\psi(s,a)\right] = \mathbb{E}_{\pi_\theta}\left[\nabla_\theta{\log\pi_\theta(a|s)}\psi(s,a)\right].
  \label{eqn:like-grad}
\end{equation}
intuitively meaning we should to increase the probability of actions judged as good by the score function $\psi$. We present some frequently used score functions in section \ref{sec:pge}.

\section{Policy Gradient Estimation}
\label{sec:pge}

In this section, we discuss some approaches to calculate the estimated policy gradient $\nabla_\theta J(\theta) = \mathbb{E}_{\pi_\theta}\left[\nabla_\theta{\log\pi_\theta(a|s)}\psi(a|s)\right]$ using different score functions to substitute $\psi$.

\paragraph{Immediate updates.} The first class of algorithms developed to update a policy directly instead of deriving it from a value function approximation were REINFORCE \cite{Williams92simplestatistical} algorithms. REINFORCE uses immediate rewards with a reinforcement baseline $b$ in its gradient steps:
\begin{equation}
  \psi_t \leftarrow r_t-b.
\end{equation}

Obviously, this does not involve temporal information. To credit temporal information $\psi_t$ needs to involve multiple time steps, meaning we need to delay updates until we have enough information to evaluate the gradient.

\paragraph{Episodic updates.} A common way of crediting temporal information in score function is to sample trajectories, also called roll-outs, and to updates using the complete roll-out. We call these approaches episodic. When we do this, we sum up the immediate updates, which we can rearrange to get
\begin{equation}
  \sum_{t=0}^h \nabla_\theta J(\theta) = \mathbb{E}_{\pi_\theta}\left[\sum_{t=0}^h \nabla_\theta{\log\pi_\theta(a|s)}\psi(s,a)\right].
  \label{eqn:egrad}
\end{equation}

Episodic REINFORCE \cite{Williams92simplestatistical}, the first approach to this concept performs roll-outs and takes only the final accumulated reward:
\begin{equation}
  \sum_{t=0}^h \nabla_\theta J(\theta) = \mathbb{E}_{\pi_\theta}\left[\sum_{t=0}^h \nabla_\theta\log\pi_\theta(a|s) \mathcal{R}^h \right].
\end{equation}

However, taking this a step further, it should be possible to attribute the reward at each step to the respective action. A very basic episodic score function that fits into our framework in \eqref{eqn:egrad} is to use the discounted rewards, similar to \eqref{eqn:acc-reward}:
\begin{equation}
  \psi_t \leftarrow \gamma^t r_t.
\end{equation}
If we only take batches of episodes starting at time $t'$, we can also use only the rewards following a chosen action $a_{t'}$ by calculating only $\sum_{t=t'}^h \nabla_\theta J(\theta)$ with $\psi_t \leftarrow \gamma^{t'} r_{t}$. As we will see, this concept is also used in value functions \eqref{eqn:v}, and \eqref{eqn:q}.

Given $r_t > 0, \forall t=0,\dots,h$, these approaches can only increase action probabilities. Obviously, we normalize to ensure $\forall s \in \mathbb{X}: \int_\mathbb{U}{\pi(a|s)da} = 1$. This means that actions can only become less probable in relation to other actions. We find that this introduces a lot of variance when learning from samples \cite{Sutton:1999:PGM:3009657.3009806}. One approach to counter the variance is to introduce a baseline
\begin{equation}
  \psi_t \leftarrow \gamma^{t} (r_{t} - b(s_t))
\end{equation}
The baseline can represent any prior knowledge or assumption about states, e.g., their mean value. Though we find that there is a specific way to use this baseline which fits surprisingly well into this framework.

\paragraph{Value functions.} Given we know the actual value, i.e. the expected accumulated reward we will get when we start in a state $s_t$, this function can be used to evaluate the performance of our policy.
\begin{equation}
	V^{\pi}(s_t) := \mathbb{E}_{\substack{s_{t+1:h} \\ a_{t:h}}}\left[\sum_{l=0}^{h} \gamma^{t+l} r_{t+l}\right].
	\label{eqn:v}
\end{equation}
Though, in general we need to approximate this function. This is the core idea of value-function approximation, but is also important in the context of policy gradients, especially for actor-critic approaches, which we discuss in \ref{sec:ac}.

Another frequently used function is the state-action value function. Instead of the expected accumulated reward starting from state $s_t$, this function gives the expected accumulated reward given an action $a_t$ is selected
\begin{equation}
	Q^{\pi}(s_t, a_t) := \mathbb{E}_{\substack{s_{t+1:h} \\ a_{t+1:h}}}\left[\sum_{l=0}^{h} \gamma^{t+l} r_{t+l}\right].
	\label{eqn:q}
\end{equation}

Approximating this function by an unbiased estimator $f_w(s_t, a_t) = \hat{Q}^\pi(s_t, a_t)$, we can also use this estimation as a score function $\psi \leftarrow \hat{Q}$. Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806} show that using this function approximation we will converge to the same local optimum. However, they also find this requires $\int_\mathbb{X}{\pi(s,a)f_w(s,a)da} = 0$ for each state. Given this assumption, the function estimator $f_w$ resembles an advantage function
\begin{equation}
	A^{\pi}(s_t, a_t) := Q^{\pi}(s_t, a_t) - V^{\pi}(s_t)
	\label{eqn:adv}
\end{equation}
as in \cite{1993b}.

Whenever we require estimating a value function for updating our policy, we can name the policy actor, and the estimated value function critic. From this observation, we can define a class of policy optimization methods called actor-critic methods, see section \ref{sec:ac}.

\section{Actor-Critic Methods} 
\label{sec:ac}

\tikzset{block/.style= {draw, rectangle, align=center, minimum height=2em, minimum width=3cm}}
\begin{figure}
  \begin{tikzpicture}[auto, node distance=2cm,>=latex']
    \node [block, name=env, minimum width=6cm] (env) {Environment};
    \node [block, above of=env, node distance=1.5cm] (critic) {Critic \\ $\hat{Q}^\pi(s,a)$};
    \node [block, above of=critic] (actor) {Actor \\ policy $\pi(a|s)$};

    \draw[densely dotted] ($(actor.north west)+(-1.5,0.6)$) rectangle ($(critic.south east)+(1.5,-0.15)$);
    \node[above of=actor, node distance=0.7cm] {Agent};

    \draw [->, align=left, swap, densely dashed] (critic.north) -- node{Policy \\ Improvement} (actor.south);
    \draw [->, pos=0.75] ($(env.north west)!0.15!(env.north)$) |- node{$s_t$} (actor.west);
    \draw [->, pos=0.75] ($(env.north west)!0.15!(env.north)$) |- node{$s_t, r_t$} (critic.west);
    \draw [->, pos=0.25] (actor.east) -| node{$a_t$} ($(env.north east)!0.15!(env.north)$);
    \draw [->, pos=0.75, swap] ($(env.north east)!0.15!(env.north)$) |- node{$a_t$} (critic.east);

    \draw [-, swap, pos=0.14] ($(env.north west)!0.15!(env.north)$) |- node{Observation} (critic.west);
    \draw [-, swap, pos=0.95] (actor.east) -| node{Action} ($(env.north east)!0.15!(env.north)$);
  \end{tikzpicture}
  \caption{An actor-critic framework comparable to Kimura et al.  \cite{Kimura1998AnAO}.} \label{fig:ac}
\end{figure}

Policy gradient methods can be described in terms of two main steps often called policy evaluation and policy improvement. For actor-critic approaches, we separate these steps from the actor component by implementing a critic. This means, the actor consists only of the policy and how to improve it given feedback from the critic, which is focused on estimating a score for the actions taken. By that concept, observations of $s_t$ from the environment are given to the actor only to decide the next action, and to the critic only to improve its function estimation. Figure \ref{fig:ac} shows the general structure of an actor-critic algorithm.

\paragraph{Critic.} The critic estimates a value function as defined in \eqref{eqn:v}. \cite{Sutton:1999:PGM:3009657.3009806}, and \cite{Konda:2003:AA:942271.942292} find that the estimation $f_w^\pi(s,a) \approx Q^\pi(s,a)$ does not affect the unbiasedness of the gradient estimate under some restrictions. Specifically, this holds for $f_w^\pi(s,a) = {\nabla_\theta \log\pi(a|s)}^T w$, thus $f_w^\pi(s,a)$ being a linear function parameterized by the vector $w$. However, as we have seen in the last section, \ref{sec:pge}, this function estimator will have a mean of zero $\int_\mathbb{U} \nabla_\theta \pi(a|s)du $, thus it estimates an advantage function \ref{eqn:adv}. However, to estimate the advantage function, we require $V^\pi(s)$. We will show how this leads to a new problem but still, this insight guarantees that function approximation does not cause divergence, and really enables recent research in reinforcement learning for continuous control problems, e.g., in humanoid robotics.

\paragraph{Actor.} The main challenge we tackle in the actor other than choosing an actual policy is how to do the policy improvement step. Though, one could argue that this step is somewhere in between the critic and the actor.

Traditionally, the improvement is often done by TD \cite{Sutton1988}, i.e., we use the temporal difference between the critic's estimations 
\begin{equation}
  r_t + \gamma \hat{V}^\pi(s_{t+1}) - \hat{V}^\pi(s_t).
\end{equation}

This is where we need to come back to the insight that from $f_w^\pi(s,a) = {\nabla_\theta \log\pi(a|s)}^T w$, it follows that $f_w^\pi(s,a)$ estimates the advantage function, and we require some form of bootstrapping for $V^\pi(s,a)$. If we use temporal difference in this context, we run into a bootstrapping problem, as \eqref{eqn:adv} subtracts $\hat{V}^\pi(s_t)$, meaning we would only learn immediate rewards \cite{Peters_IICHR_2003}. This would render the process biased. Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806} and Konda et al. \cite{NIPS1999_1786} suggest estimating an action value function as in \eqref{eqn:q}. We can approximate this $f_w^\pi$ by least-squares optimization over multiple $\hat{Q}^\pi(s,a)$ obtained from roll-outs. However, Peters et al. \cite{4863} find that this approximation is highly reliant on the distribution of the training data. This comes from the realization, that we use only a subspace of the actual action value function in $f_w^\pi$, which is only a state value function. One can compare this to approximating a parabola by a line, whereby the approximation changes wildly depending on which part of the parabola is in the training data. An approach to solve this bootstrapping problem is to rewrite the Bellman Equation using \eqref{eqn:adv} and \eqref{eqn:q} with $A^\pi(s,a) = f_w^\pi(s,a)$, $V^\pi(s) = \phi(s)^T v$, and a zero-mean error term $\epsilon(s_t,a_t,s_{t+1})$, we get
\begin{align}
  Q^\pi(s,a) = A^\pi(s,a) + V^\pi(s) = r(s,a) + \gamma \int_\mathbb{X} p(s'|s,a)V^\pi(s')ds' \\
  \nabla_\theta \log \pi(a_t|s_t)^T w + \phi(s_t)^T v = r(s_t,a_t) + \gamma \phi(s_{t+1})^T v + \epsilon(s_t,a_t,s_{t+1})
\end{align}
which gives involves only linear equations to solve. \cite{4863}

With these insights in mind, \ref{sec:natural} presents the natural gradient, a refined type of gradient which has a convenient fit in the actor-critic setting we just established.

\section{Natural Gradient}
\label{sec:natural}

Natural gradients were at first proposed for use in supervised learning settings by Amari et al. \cite{Amari:1998:NGW:287476.287477}, but have been shown to be effective in reinforcement learning by Kakade \cite{Kakade:2001} and Peters et al. \cite{4863}.

When using normal gradient steps, we find that steps can become very small when a plateau is reached. This can will drastically slow down the learning process, and in the worst case cause some algorithms to prematurely. However, we can use some additional information to refine the gradient. Figure \ref{fig:nat-grad-adv} shows an example by Peters et al. \cite{Peters_IICHR_2003} that gives a visual intuition about the difference between 'vanilla' and natural policy gradients.

\begin{figure}
  \includegraphics[width=\textwidth]{nat-grad-adv}
  \caption{An experiment showing where the natural gradient has a great advantage. \cite{Peters_IICHR_2003} }\label{fig:nat-grad-adv}
\end{figure}

Using the Fisher information matrix $F_\theta$, and the gradient estimate we discussed in section \ref{sec:pge} gives us the definition
\begin{equation}
  \widetilde{\nabla}_\theta J(\theta) := F^{-1}_\theta \nabla_\theta J(\theta)
  \label{eqn:nat-grad}
\end{equation}
of the natural gradient. The Fisher information matrix represents the certainty we have on our estimate of the gradient and is defined as the covariance of the log likelihood function of a trajectory $\rho_{\pi_\theta}^h$, which as \cite{4863} show can be written as 
\begin{equation}
  F_\theta = \int_\mathbb{X} p(s|\theta) \int_\mathbb{U} \pi_\theta(a|s) \nabla_\theta \log{\pi_\theta(a|s)} \nabla_\theta \log{\pi_\theta(a|s)}^T dads.
  \label{eqn:F}
\end{equation}
When we recall the definition \eqref{eqn:like-grad} of likelihood-ratio gradients, we see that setting $\psi$ as $\nabla_\theta \log \pi(a|s)^T w$, we get
\begin{equation}
  \nabla_\theta J(\theta) = \int_\mathbb{X} {p(s|\theta) \int_\mathbb{U} \pi_\theta(a|s) \nabla_\theta \log \pi(a|s) {\nabla_\theta \log \pi(a|s)}^T dads}\ w := F_\theta w.
  \label{eqn:J-equals-F}
\end{equation}

From \eqref{eqn:nat-grad}, \eqref{eqn:F}, and \eqref{eqn:J-equals-F}, it follows that
\begin{equation}
  \widetilde{\nabla}_\theta J(\theta) = F^{-1}_\theta \nabla_\theta J(\theta) = F_\theta^{-1} F_\theta w = w.
\end{equation}
Thus, this approach does not require an actual estimate of the Fisher information matrix, but only an estimate of $w$, with the update step according to $\theta_{k+1} = \theta_k + \alpha_k w$.

Peters et al. \cite{4863} present this idea and suggest LSTD($\lambda$)-Q, a version of least-squares temporal difference learning  \cite{Boyan:1999:LTD:645528.657618}, as well as episodic natural actor-critic.

Note that this approach abuses some specific conveniences about the problem setup to find an elegant solution. However, this is not the only way to approach this problem. In this paper, we focus on one presentation to present the core idea, but the field of policy gradient optimization is surprisingly resourceful. In the conclusion \ref{sec:outro}, we attempt to show the versatility and relevance in current research, and give some directions for further reading.

\section{Conclusion}
\label{sec:outro}

In this paper, we have introduced policy gradient methods as a class of reinforcement learning algorithms. We show why policy gradient methods are effective in these environments, and we give some intuitions about the concept. Further, we show the core elements of policy gradient methods, and we discuss some intricacies that they bring, and some research development in the attempts of improving the efficiency and stability of policy gradients.

There are many possibilities for improvement as it can be seen as a very modularized and flexible. We find that the theoretical basis for policy gradient focused approaches reaches far back with some fundamental algorithms like Levenberg-Marquardt \cite{levenbergmarquardt1978}, and the first presentation of REINFORCE \cite{Williams92simplestatistical}. Policy gradient methods were shown to be increasingly potent by authors such as Sutton \cite{Sutton:1999:PGM:3009657.3009806}, and still provide space for a lot of new work being published. Algorithms such as the aforementioned Natural Actor-Critic \cite{4863} have been showing some of their potentials, but even so potential areas of improvement. Schulman et al. \cite{DBLP:journals/corr/SchulmanMLJA15} present a form of advantage estimation using $\gamma$ as a parameter for bias-variance trade-off. Fellows et al. \cite{fellows:icml18} use Fourier analysis to analytically improve policy gradients. Abdolmaleki et al. \cite{abdolmaleki2018maximum} introduce MPO, an off-policy algorithm using expectation maximization, related to the approach of TRPO \cite{pmlr-v37-schulman15}. Yet a different approach is given by Zhang et al. \cite{DBLP:journals/corr/abs-1808-03030}, who formulate policy optimization as a Wasserstein gradient flow problem, which moves into a Wasserstein manifold in the distribution space. This is comparable to approaches by Abdolmaleki et al. \cite{NIPS2015_5672}, which introduces a bound on the Kullback-Leibler divergence during policy search. From the developments in recent research, it is fair to say that policy gradient methods play a major role in reinforcement learning.

\bibliographystyle{spmpsci}
\bibliography{bibliography}

\end{document}

### 2019-02-27 pge-mattis.tex
TODOs and minor fixes

@@
 
-\author {Mattis Manfred K{\"a}mmerer \and Mahdi Enan \and Simon Kiefhaber}
+\author {Mattis Manfred K{\"a}mmerer}
 
-\institute{Mattis Manfred K{\"a}mmerer \and Mahdi Enan \and Simon Kiefhaber \at TU Darmstadt}
+\institute{TU Darmstadt}
 
@@
 \begin{abstract}
+
+TODO
+
 Policy gradient methods use a policy model to decide which actions maximize the expected accumulated reward received in a given environment.
@@
 
-\keywords{reinforcement learning \and policy gradients \and actor-critic \and advantage estimation \and continuous control }
+\keywords{reinforcement learning \and policy gradients \and actor-critic \and natural gradient }
 \end{abstract}
@@
 
-Introductions to reinforcement learning often start by introducing value-based methods.
-They aim to gain the maximum accumulated reward in a Markov Decision Process (MDP).
-Value-based methods estimate the expected value of each action in each state, approximating a function that ranks actions in a given state, which is then used to derive a policy.
-Though this works well for problems with countable actions and states, its performance is not stable on continuous control problems.
-Even with optimal hyperparameters, the outcomes vary between separate trainings.
-\cite{DBLP:journals/corr/MetzIJD17} In constract to value-based methods, the other main class of model-free reinforcement learning algorithms is called policy-based methods, or policy gradient methods.
-Research has shown, that policy gradient methods can reach better performance in these environments.
-\cite{pmlr-v37-schulman15} Using a parameterized (often stochastic) policy to decide the next action, they can easily incorporate prior domain knowledge, and reuse some of the value-based approaches for improved performance, but require a lot of configuration to produce an effective agent for a specific environment.
-Also, they frequently require on-policy training and often converge slower in discrete settings.
+Policy gradient methods are approaches to maximize the expected return in a Markov Decision Process (MDP).
+Using a parameterized policy to decide the next action, they can easily incorporate prior domain knowledge, but require a lot of configuration to produce an effective agent for a specific environment.
+Also, they frequently require on-policy training and often converge slower, but can guaranteed find a local optimum given some constraints on the hyperparameters.
 Options to model a policy gradient method are not only given by the policy model itself, but also in the way a policy improvement is computed.
-\cite{DBLP:journals/corr/SchulmanMLJA15} In this case, improvement means a step in the parameter space such that the policy under the new parameters will on average perform better than the old policy.
+\cite{DBLP:journals/corr/SchulmanMLJA15} In this case, improvement means a step in the parameter space such that the policy under the new parameters will on average perform better than the old policy, i.e.
+improve its expected return.
 Policy gradient estimation is the term we use to describe the process of computing the next step in parameter space.
-The goal is to estimate a function that predicts how the policy should be changed to maximize the accumulated reward of an agent in an MDP.
+The goal is to estimate the gradient of the policy with respect to the expected return.
 Since this is the core problem of policy gradient methods, it is also the main topic of this paper.
@@
 
-\paragraph{Definitions.} We define states $s \in \mathbb{X}$, actions $a \in \mathbb{U}$, and rewards $r \in \mathbb{R}$.
-A trajectory $\rho_{\pi_\theta}^h := (s_0, a_0, $ $s_1, a_1, \dots, s_h, a_h)$ is generated by drawing $s_0 \sim p(x_0)$ according to the distribution over initial states $p(x_0)$, and successively sampling $a_t \sim \pi_\theta(a_t|s_t)$ according to the policy $\pi_\theta$, and $s_{t+1} \sim p(s_{t+1}|s_t,a_t)$ until the horizon $h$ is reached.
+\paragraph{Definitions.} We define states $s \in \mathbb{S}$, actions $a \in \mathbb{A}$, and rewards $r \in \mathbb{R}$.
+A trajectory $\tau_{\pi_\theta}^h := (s_0, a_0, $ $s_1, a_1, \dots, s_h, a_h)$ is generated by drawing $s_0 \sim p(x_0)$ according to the distribution over initial states $p(x_0)$, and successively sampling $a_t \sim \pi_\theta(a_t|s_t)$ according to the policy $\pi_\theta$, and $s_{t+1} \sim p(s_{t+1}|s_t,a_t)$ until the horizon $h$ is reached.
 At each time step, we receive a reward according to $r_t = r(s_t, a_t, s_{t+1})$.
@@
 
-We assume a Markov Decision Process (MDP), meaning the proability distribution of the next states is independent of past states $s_{0:t-1}$ given the present state $s_t$ and action $a_t$, 
+We assume a Markov Decision Process (MDP), meaning the probability distribution of the next states is independent of past states $s_{0:t-1}$ given the present state $s_t$ and action $a_t$, 
 \begin{equation}
@@
 
-\paragraph{Policy gradient.} Our goal with respect to episodes is to maximize the expectation of the total reward $\mathbb{E}_{s\sim p(s|\theta)}\left[\mathcal{R}^\infty\right]$.
-The total reward in the horizon $h$ for the non-discounted problem is $\sum_{t=0}^{h} r_{t}$.
-For most cases, a discount factor $\gamma \in [0,1)$ is added to ensure that this sum is finite for the infinite horizon $h\to\infty$.
-Intuitively, this reflects the idea that the relevance of later actions declines, and reduces variance during learning.
+\paragraph{Policy gradient.} Our goal with respect to episodes is to maximize the expectation of the total reward $\mathbb{E}_{s\sim p(s|\theta)}\left[\mathcal{R}^\infty\right]$, also called expected return.
+We define $p(s|\theta)$ as the probability The total reward in the horizon $h$ is $\sum_{t=0}^{h} r_{t}$.
+We introduce a discount factor $\gamma \in [0,1)$ to trade-off bias ($\gamma\to0$) versus variance ($\gamma\to1$) with a normalization factor of $1-\gamma$, which is often omitted though strongly recommended.
+Intuitively, this reflects the idea that the relevance of later actions declines, which puts more weight on less actions, thus the increased bias for reduced variance.
 The discounted total reward is
@@
 \begin{equation}
-  \mathcal{R}^h := \sum_{t=0}^{h} \gamma^t r_t.
+  \mathcal{R}_0^h = \mathcal{R}^h := (1-\gamma) \sum_{t=0}^{h} \gamma^t r_t.
   \label{eqn:acc-reward}
@@
 
-Since we have only empirical knowledge of the performance of the policy, we need to approximate an optimal policy by estimating a gradient.
-Thus, we search $\nabla_\theta J(\theta) := \nabla_\theta \mathbb{E}_{s\sim p(s|\theta)}\left[\psi(s,a)\right]$, to make a policy gradient step according to $\theta_{k+1} = \theta_k + \alpha_k \nabla_\theta J(\theta)$, where $\psi$ denotes some score function, see section \ref{sec:pge}, and $\alpha_k$ denotes a learning rate.
+Since we have only limited knowledge of the performance of the policy, we need to approximate an optimal policy by estimating a gradient.
+Thus, we search $\nabla_\theta J(\theta) := \nabla_\theta \mathbb{E}_{p(\tau|\pi)}\left[\psi(\tau)\right]$, to make a policy gradient step according to $\theta_{k+1} = \theta_k + \alpha_k \nabla_\theta J(\theta)$, where $\psi$ denotes some score function, see section \ref{sec:pge}, and $\alpha_k$ denotes a learning rate.
 It is shown that for $\lim_{k\to\infty}\alpha_k = 0$, and $\sum_{k=0}^\infty \alpha_k$ we are guaranteed to converge to a local optimum \cite{Sutton:1999:PGM:3009657.3009806}.
 
-\paragraph{Likelihood-ratio gradients.} Using the definition of the expectation, we derive
-\begin{align*}
-  \nabla_\theta \mathbb{E}_{s\sim p(s|\theta)}\left[\psi(s,a)\right] &= \nabla_\theta \int_\mathbb{X}{p(s|\theta)\int_\mathbb{U}{\pi_\theta(a|s)\psi(s,a) da}ds} \\
-  &= \int_\mathbb{X}{p(s|\theta)\int_\mathbb{U}{\nabla_\theta {\pi_\theta(a|s)} \psi(s,a) da}ds} \\
-  &= \int_\mathbb{X}{p(s|\theta)\int_\mathbb{U}{\pi_\theta(a|s)\nabla_\theta{\log\pi_\theta(a|s)} \psi(s,a) da}ds} .
-\end{align*}
+\paragraph{Finite-difference gradients.} TODO
+
+\paragraph{Likelihood-ratio gradients.} TODO
+
+$P^\pi_{i,j}=P^\pi(s_j|s_i)$ Probabilities to get to $s_j$ from $s_i$
+
+$r^\pi_i=R^\pi(s_i)$ Rewards of states
+
+$\mu_0=p(s_i)$ Initial state probabilities
+
+$\mu_\theta = \mu_0 + \gamma P^T \mu_\theta$
+
+$(I - \gamma P^T)\mu_\theta = \mu_0 \implies \mu_\theta^T = \mu_0(I-\gamma P)^{-1}$
+
+$J(\theta) = \mu_0^T \nabla_\theta V^\pi$ s.t. $V^\pi = r^\pi = \gamma P^\pi V^\pi$
+
+$\nabla_\theta J(\theta) = \mu_0^T \nabla_\theta V^\pi$
+
+$\nabla_\theta J(\theta) = \mu_0^T (I - \gamma P^\theta)^{-1} (\nabla_\theta r + \gamma (\nabla_\theta P^\theta) V^\theta) = \mu_\theta^T (\nabla_\theta r^\theta + \gamma (\nabla_\theta P^\theta) V^\theta)$
+
+$r_i^\theta = \int_\mathbb{A} r(s,a)\pi_\theta(a|s) da$
+
+$\nabla_\theta J(\theta) = \sum \mu_\theta(s) \nabla_\theta\pi(a|s) Q(s,a)$
+
+LR trick
+
+$ = \sum \mu_\theta(s) \pi(a|s) \nabla_\theta\log\pi(a|s) Q(s,a)$
+
 Where we used $\nabla_\theta \pi_\theta(a|s) = \pi_\theta(a|s)\nabla_\theta\log\pi_\theta(a|s)$, obtained from the likelihood ratio, which gives us the likelihood-ratio gradient,
 \begin{equation}
-  \nabla_\theta \mathbb{E}_{\pi_\theta}\left[\psi(s,a)\right] = \mathbb{E}_{\pi_\theta}\left[\nabla_\theta{\log\pi_\theta(a|s)}\psi(s,a)\right].
+  \nabla_\theta \mathbb{E}_{\pi_\theta}\left[\psi(\tau)\right] = \mathbb{E}_{\pi_\theta}\left[\nabla_\theta{\log\pi_\theta(a|s)}\psi(\tau)\right].
   \label{eqn:like-grad}
@@
 \paragraph{Immediate updates.} The first class of algorithms developed to update a policy directly instead of deriving it from a value function approximation were REINFORCE \cite{Williams92simplestatistical} algorithms.
-REINFORCE uses immediate rewards with a reinforcement baseline $b$ in its gradient steps:
+REINFORCE uses immediate updates with a baseline $b$ in its gradient steps:
 \begin{equation}
@@
 \begin{equation}
-  \sum_{t=0}^h \nabla_\theta J(\theta) = \mathbb{E}_{\pi_\theta}\left[\sum_{t=0}^h \nabla_\theta{\log\pi_\theta(a|s)}\psi(s,a)\right].
+  \sum_{t=0}^h \nabla_\theta J(\theta) = \mathbb{E}_{\pi_\theta}\left[\sum_{t=0}^h \nabla_\theta{\log\pi_\theta(a|s)}\psi(\tau)\right].
   \label{eqn:egrad}
@@
 
-Episodic REINFORCE \cite{Williams92simplestatistical}, the first approach to this concept performs roll-outs and takes only the final accumulated reward:
+Episodic REINFORCE \cite{Williams92simplestatistical}, the first approach to this concept performs roll-outs and takes only the final return:
 \begin{equation}
-  \sum_{t=0}^h \nabla_\theta J(\theta) = \mathbb{E}_{\pi_\theta}\left[\sum_{t=0}^h \nabla_\theta\log\pi_\theta(a|s) \mathcal{R}^h \right].
+  \nabla_\theta J(\theta) \approx \frac{1}{h} \sum_{t=0}^h \nabla_\theta\log\pi_\theta(a|s) \mathcal{R}_t^h.
 \end{equation}
-
+This is sometimes also called Monte-Carlo gradient estimation.
 However, taking this a step further, it should be possible to attribute the reward at each step to the respective action.
@@
 Given $r_t > 0, \forall t=0,\dots,h$, these approaches can only increase action probabilities.
-Obviously, we normalize to ensure $\forall s \in \mathbb{X}: \int_\mathbb{U}{\pi(a|s)da} = 1$.
+Obviously, we normalize to ensure $\forall s \in \mathbb{S}: \int_\mathbb{A}{\pi(a|s)da} = 1$.
 This means that actions can only become less probable in relation to other actions.
@@
 \begin{equation}
-	V^{\pi}(s_t) := \mathbb{E}_{\substack{s_{t+1:h} \\ a_{t:h}}}\left[\sum_{l=0}^{h} \gamma^{t+l} r_{t+l}\right].
+	V^{\pi}(s_t) := \mathbb{E}_{\substack{s_{t+1:h} \\ a_{t:h}}}\left[\mathcal{R}_t^h\right].
 	\label{eqn:v}
 \end{equation}
-Though, in general we need to approximate this function.
-This is the core idea of value-function approximation, but is also important in the context of policy gradients, especially for actor-critic approaches, which we discuss in \ref{sec:ac}.
+
+In general, we need to approximate this function.
+Obviously, REINFORCE does this approximation by a simple Monte-Carlo sample.
+However, there is a formulation of policy gradient methods that implements this approximation problem explicitly called actor-critic, which we discuss in \ref{sec:ac}.
 
@@
 \begin{equation}
-	Q^{\pi}(s_t, a_t) := \mathbb{E}_{\substack{s_{t+1:h} \\ a_{t+1:h}}}\left[\sum_{l=0}^{h} \gamma^{t+l} r_{t+l}\right].
+	Q^{\pi}(s_t, a_t) := \mathbb{E}_{\substack{s_{t+1:h} \\ a_{t+1:h}}}\left[\mathcal{R}_t^h\right].
 	\label{eqn:q}
@@
 \cite{Sutton:1999:PGM:3009657.3009806} show that using this function approximation we will converge to the same local optimum.
-However, they also find this requires $\int_\mathbb{X}{\pi(s,a)f_w(s,a)da} = 0$ for each state.
+However, they also find this requires $\int_\mathbb{S}{\pi(s,a)f_w(s,a)da} = 0$ for each state.
 Given this assumption, the function estimator $f_w$ resembles an advantage function
@@
 Whenever we require estimating a value function for updating our policy, we can name the policy actor, and the estimated value function critic.
-From this observation, we can define a class of policy optimization methods called actor-critic methods, see section \ref{sec:ac}.
+From this observation, we define a class of policy optimization methods called actor-critic methods in section \ref{sec:ac}.
 
@@
 For actor-critic approaches, we separate these steps from the actor component by implementing a critic.
-This means, the actor consists only of the policy and how to improve it given feedback from the critic, which is focused on estimating a score for the actions taken.
-By that concept, observations of $s_t$ from the environment are given to the actor only to decide the next action, and to the critic only to improve its function estimation.
+This means, the actor consists only of the policy, while the critic is focused on estimating a score for the actions taken.
+By that concept, observations of $s_t$ from the environment are given to the actor only to decide the next action, and to the critic only to improve its function estimation with the respective $r_t$.
 Figure \ref{fig:ac} shows the general structure of an actor-critic algorithm.
+
+\paragraph{Actor.} TODO
 
@@
 Specifically, this holds for $f_w^\pi(s,a) = {\nabla_\theta \log\pi(a|s)}^T w$, thus $f_w^\pi(s,a)$ being a linear function parameterized by the vector $w$.
-However, as we have seen in the last section, \ref{sec:pge}, this function estimator will have a mean of zero $\int_\mathbb{U} \nabla_\theta \pi(a|s)du $, thus it estimates an advantage function \ref{eqn:adv}.
+However, as we have seen in the last section, \ref{sec:pge}, this function estimator will have a mean of zero $\int_\mathbb{A} \nabla_\theta \pi(a|s)du $, thus it estimates an advantage function \ref{eqn:adv}.
 However, to estimate the advantage function, we require $V^\pi(s)$.
@@
 
-\paragraph{Actor.} The main challenge we tackle in the actor other than choosing an actual policy is how to do the policy improvement step.
-Though, one could argue that this step is somewhere in between the critic and the actor.
-
-Traditionally, the improvement is often done by TD \cite{Sutton1988}, i.e., we use the temporal difference between the critic's estimations 
+Traditionally, the improvement is often done by Monte-Carlo sampling as in REINFORCE, or using temporal difference (TD) \cite{Sutton1988}, i.e., we use the temporal difference between the critic's estimations
 \begin{equation}
@@
 
-This is where we need to come back to the insight that from $f_w^\pi(s,a) = {\nabla_\theta \log\pi(a|s)}^T w$, it follows that $f_w^\pi(s,a)$ estimates the advantage function, and we require some form of bootstrapping for $V^\pi(s,a)$.
-If we use temporal difference in this context, we run into a bootstrapping problem, as \eqref{eqn:adv} subtracts $\hat{V}^\pi(s_t)$, meaning we would only learn immediate rewards \cite{Peters_IICHR_2003}.
+This is where we need to come back to the insight that from $f_w^\pi(s,a) = {\nabla_\theta \log\pi(a|s)}^T w$, it follows that $f_w^\pi(s,a)$ estimates the advantage function, so we are bootstrapping for $V^\pi(s,a)$.
+If we use temporal difference in this context, we run into a problem, as \eqref{eqn:adv} subtracts $\hat{V}^\pi(s_t)$, meaning we would only learn immediate rewards \cite{Peters_IICHR_2003}.
 This would render the process biased.
@@
 \begin{align}
-  Q^\pi(s,a) = A^\pi(s,a) + V^\pi(s) = r(s,a) + \gamma \int_\mathbb{X} p(s'|s,a)V^\pi(s')ds' \\
+  Q^\pi(s,a) = A^\pi(s,a) + V^\pi(s) = r(s,a) + \gamma \int_\mathbb{S} p(s'|s,a)V^\pi(s')ds' \\
   \nabla_\theta \log \pi(a_t|s_t)^T w + \phi(s_t)^T v = r(s_t,a_t) + \gamma \phi(s_{t+1})^T v + \epsilon(s_t,a_t,s_{t+1})
@@
 of the natural gradient.
-The Fisher information matrix represents the certainty we have on our estimate of the gradient and is defined as the covariance of the log likelihood function of a trajectory $\rho_{\pi_\theta}^h$, which as \cite{4863} show can be written as
+The Fisher information matrix represents the certainty we have on our estimate of the gradient and is defined as the covariance of the log likelihood function of a trajectory $\tau_{\pi_\theta}^h$, which as \cite{4863} show can be written as
 \begin{equation}
-  F_\theta = \int_\mathbb{X} p(s|\theta) \int_\mathbb{U} \pi_\theta(a|s) \nabla_\theta \log{\pi_\theta(a|s)} \nabla_\theta \log{\pi_\theta(a|s)}^T dads.
+  F_\theta = \int_\mathbb{S} p(s|\theta) \int_\mathbb{A} \pi_\theta(a|s) \nabla_\theta \log{\pi_\theta(a|s)} \nabla_\theta \log{\pi_\theta(a|s)}^T dads.
   \label{eqn:F}
@@
 \begin{equation}
-  \nabla_\theta J(\theta) = \int_\mathbb{X} {p(s|\theta) \int_\mathbb{U} \pi_\theta(a|s) \nabla_\theta \log \pi(a|s) {\nabla_\theta \log \pi(a|s)}^T dads}\ w := F_\theta w.
+  \nabla_\theta J(\theta) = \int_\mathbb{S} {p(s|\theta) \int_\mathbb{A} \pi_\theta(a|s) \nabla_\theta \log \pi(a|s) {\nabla_\theta \log \pi(a|s)}^T dads}\ w := F_\theta w.
   \label{eqn:J-equals-F}
@@
 However, this is not the only way to approach this problem.
-In this paper, we focus on one presentation to present the core idea, but the field of policy gradient optimization is surprisingly resourceful.
-In the conclusion \ref{sec:outro}, we attempt to show the versatility and relevance in current research, and give some directions for further reading.
+In this paper, we focus on one presentation to present the core idea, but the field of policy gradient optimization is very resourceful.
+TODO In the conclusion \ref{sec:outro}, we attempt to show the versatility and relevance in current research, and give some directions for further reading.
 
@@
 \label{sec:outro}
+
+TODO less references and stuff...
 

### 2019-03-03 pge-mattis.tex
Sentences in separate lines

@@
 Also, they frequently require on-policy training and often converge slower, but can guaranteed find a local optimum given some constraints on the hyperparameters. 
-Options to model a policy gradient method are not only given by the policy model itself, but also in the way a policy improvement is computed.
-\cite{DBLP:journals/corr/SchulmanMLJA15} In this case, improvement means a step in the parameter space such that the policy under the new parameters will on average perform better than the old policy, i.e.
-improve its expected return.
+Options to model a policy gradient method are not only given by the policy model itself, but also in the way a policy improvement is computed \cite{DBLP:journals/corr/SchulmanMLJA15}. 
+In this case, improvement means a step in the parameter space such that the policy under the new parameters will on average perform better than the old policy, i.e. improve its expected return. 
 Policy gradient estimation is the term we use to describe the process of computing the next step in parameter space. 
@@
 
-\paragraph{Definitions.} We define states $s \in \mathbb{S}$, actions $a \in \mathbb{A}$, and rewards $r \in \mathbb{R}$.
+\paragraph{Definitions.} 
+We define states $s \in \mathbb{S}$, actions $a \in \mathbb{A}$, and rewards $r \in \mathbb{R}$. 
 A trajectory $\tau_{\pi_\theta}^h := (s_0, a_0, $ $s_1, a_1, \dots, s_h, a_h)$ is generated by drawing $s_0 \sim p(x_0)$ according to the distribution over initial states $p(x_0)$, and successively sampling $a_t \sim \pi_\theta(a_t|s_t)$ according to the policy $\pi_\theta$, and $s_{t+1} \sim p(s_{t+1}|s_t,a_t)$ until the horizon $h$ is reached.
@@
 
-\paragraph{Policy gradient.} Our goal with respect to episodes is to maximize the expectation of the total reward $\mathbb{E}_{s\sim p(s|\theta)}\left[\mathcal{R}^\infty\right]$, also called expected return.
-We define $p(s|\theta)$ as the probability The total reward in the horizon $h$ is $\sum_{t=0}^{h} r_{t}$.
+\paragraph{Policy gradient.} 
+Our goal with respect to episodes is to maximize the expectation of the total reward $\mathbb{E}_{s\sim p(s|\theta)}\left[\mathcal{R}^\infty\right]$, also called expected return. 
+We define $p(s|\theta)$ as the probability TODO
+The total reward in the horizon $h$ is $\sum_{t=0}^{h} r_{t}$. 
 We introduce a discount factor $\gamma \in [0,1)$ to trade-off bias ($\gamma\to0$) versus variance ($\gamma\to1$) with a normalization factor of $1-\gamma$, which is often omitted though strongly recommended.
@@
 
-\paragraph{Immediate updates.} The first class of algorithms developed to update a policy directly instead of deriving it from a value function approximation were REINFORCE \cite{Williams92simplestatistical} algorithms.
+\paragraph{Immediate updates.} 
+The first class of algorithms developed to update a policy directly instead of deriving it from a value function approximation were REINFORCE \cite{Williams92simplestatistical} algorithms.
 REINFORCE uses immediate updates with a baseline $b$ in its gradient steps:
@@
 
-\paragraph{Episodic updates.} A common way of crediting temporal information in score function is to sample trajectories, also called roll-outs, and to updates using the complete roll-out.
+\paragraph{Episodic updates.} 
+A common way of crediting temporal information in score function is to sample trajectories, also called roll-outs, and to updates using the complete roll-out. 
 We call these approaches episodic. 
@@
 Obviously, we normalize to ensure $\forall s \in \mathbb{S}: \int_\mathbb{A}{\pi(a|s)da} = 1$. 
-This means that actions can only become less probable in relation to other actions.
+This means actions can only become less probable in relation to other actions. 
 We find that this introduces a lot of variance when learning from samples \cite{Sutton:1999:PGM:3009657.3009806}. 
@@
 
-\paragraph{Value functions.} Given we know the actual value, i.e.
-the expected accumulated reward we will get when we start in a state $s_t$, this function can be used to evaluate the performance of our policy.
+\paragraph{Value functions.} 
+Given we know the actual value, i.e. the expected accumulated reward we will get when we start in a state $s_t$, this function can be used to evaluate the performance of our policy.
 \begin{equation}
@@
 
-\paragraph{Actor.} TODO
+\paragraph{Actor.} 
+TODO
 
-\paragraph{Critic.} The critic estimates a value function as defined in \eqref{eqn:v}.
+\paragraph{Critic.} 
+The critic estimates a value function as defined in \eqref{eqn:v}. 
 \cite{Sutton:1999:PGM:3009657.3009806}, and \cite{Konda:2003:AA:942271.942292} find that the estimation $f_w^\pi(s,a) \approx Q^\pi(s,a)$ does not affect the unbiasedness of the gradient estimate under some restrictions.
@@
 This would render the process biased. 
-Sutton et al.
-\cite{Sutton:1999:PGM:3009657.3009806} and Konda et al.
-\cite{NIPS1999_1786} suggest estimating an action value function as in \eqref{eqn:q}.
+Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806} and Konda et al. \cite{NIPS1999_1786} suggest estimating an action value function as in \eqref{eqn:q}. 
 We can approximate this $f_w^\pi$ by least-squares optimization over multiple $\hat{Q}^\pi(s,a)$ obtained from roll-outs. 
-However, Peters et al.
-\cite{4863} find that this approximation is highly reliant on the distribution of the training data.
+However, Peters et al. \cite{4863} find that this approximation is highly reliant on the distribution of the training data. 
 This comes from the realization, that we use only a subspace of the actual action value function in $f_w^\pi$, which is only a state value function. 
@@
 However, we can use some additional information to refine the gradient. 
-Figure \ref{fig:nat-grad-adv} shows an example by Peters et al.
-\cite{Peters_IICHR_2003} that gives a visual intuition about the difference between 'vanilla' and natural policy gradients.
+Figure \ref{fig:nat-grad-adv} shows an example by Peters et al. \cite{Peters_IICHR_2003} that gives a visual intuition about the difference between 'vanilla' and natural policy gradients.
 
@@
 In this paper, we focus on one presentation to present the core idea, but the field of policy gradient optimization is very resourceful. 
-TODO In the conclusion \ref{sec:outro}, we attempt to show the versatility and relevance in current research, and give some directions for further reading.
+TODO 
+In the conclusion \ref{sec:outro}, we attempt to show the versatility and relevance in current research, and give some directions for further reading.
 
@@
 Algorithms such as the aforementioned Natural Actor-Critic \cite{4863} have been showing some of their potentials, but even so potential areas of improvement. 
-Schulman et al.
-\cite{DBLP:journals/corr/SchulmanMLJA15} present a form of advantage estimation using $\gamma$ as a parameter for bias-variance trade-off.
-Fellows et al.
-\cite{fellows:icml18} use Fourier analysis to analytically improve policy gradients.
-Abdolmaleki et al.
-\cite{abdolmaleki2018maximum} introduce MPO, an off-policy algorithm using expectation maximization, related to the approach of TRPO \cite{pmlr-v37-schulman15}.
+Schulman et al. \cite{DBLP:journals/corr/SchulmanMLJA15} present a form of advantage estimation using $\gamma$ as a parameter for bias-variance trade-off. 
+Fellows et al. \cite{fellows:icml18} use Fourier analysis to analytically improve policy gradients. 
+Abdolmaleki et al. \cite{abdolmaleki2018maximum} introduce MPO, an off-policy algorithm using expectation maximization, related to the approach of TRPO \cite{pmlr-v37-schulman15}. 
 Yet a different approach is given by Zhang et al.
 \cite{DBLP:journals/corr/abs-1808-03030}, who formulate policy optimization as a Wasserstein gradient flow problem, which moves into a Wasserstein manifold in the distribution space.
-This is comparable to approaches by Abdolmaleki et al.
-\cite{NIPS2015_5672}, which introduces a bound on the Kullback-Leibler divergence during policy search.
+This is comparable to approaches by Abdolmaleki et al. \cite{NIPS2015_5672}, which introduces a bound on the Kullback-Leibler divergence during policy search. 
 From the developments in recent research, it is fair to say that policy gradient methods play a major role in reinforcement learning.

### 2019-03-03 pge-mattis.tex
Add more to LR gradient derivation

@@
 We define states $s \in \mathbb{S}$, actions $a \in \mathbb{A}$, and rewards $r \in \mathbb{R}$. 
+For discrete problems, we assume $\mathbb{S} \equiv \{0,1,2,\ldots\}$ and $\mathbb{A} \equiv \{0,1,2,\ldots\}$ without loss of generality.
 A trajectory $\tau_{\pi_\theta}^h := (s_0, a_0, $ $s_1, a_1, \dots, s_h, a_h)$ is generated by drawing $s_0 \sim p(x_0)$ according to the distribution over initial states $p(x_0)$, and successively sampling $a_t \sim \pi_\theta(a_t|s_t)$ according to the policy $\pi_\theta$, and $s_{t+1} \sim p(s_{t+1}|s_t,a_t)$ until the horizon $h$ is reached.
@@
 \paragraph{Policy gradient.} 
-Our goal with respect to episodes is to maximize the expectation of the total reward $\mathbb{E}_{s\sim p(s|\theta)}\left[\mathcal{R}^\infty\right]$, also called expected return. 
-We define $p(s|\theta)$ as the probability TODO
+Our goal with respect to episodes is to maximize the expectation of the total reward, also called expected return. 
 The total reward in the horizon $h$ is $\sum_{t=0}^{h} r_{t}$. 
@@
 \begin{equation}
-  \mathcal{R}_0^h = \mathcal{R}^h := (1-\gamma) \sum_{t=0}^{h} \gamma^t r_t.
+  \mathcal{R}_0^h = \mathcal{R}^\tau := (1-\gamma) \sum_{t=0}^{h} \gamma^t r_t.
   \label{eqn:acc-reward}
@@
 Since we have only limited knowledge of the performance of the policy, we need to approximate an optimal policy by estimating a gradient. 
-Thus, we search $\nabla_\theta J(\theta) := \nabla_\theta \mathbb{E}_{p(\tau|\pi)}\left[\psi(\tau)\right]$, to make a policy gradient step according to $\theta_{k+1} = \theta_k + \alpha_k \nabla_\theta J(\theta)$, where $\psi$ denotes some score function, see section \ref{sec:pge}, and $\alpha_k$ denotes a learning rate.
+Thus, we search $\nabla_\theta J(\theta) := \nabla_\theta \mathbb{E}_{p(\tau)}\left[\mathcal{R}^\tau\right]$, to make a policy gradient step according to $\theta_{k+1} = \theta_k + \alpha_k \nabla_\theta J(\theta)$, where $\alpha_k$ denotes a learning rate.
 It is shown that for $\lim_{k\to\infty}\alpha_k = 0$, and $\sum_{k=0}^\infty \alpha_k$ we are guaranteed to converge to a local optimum \cite{Sutton:1999:PGM:3009657.3009806}.
@@
 
-\paragraph{Likelihood-ratio gradients.} TODO
+Since finite-difference gradients suffer from the curse of dimensionality, they are too inefficient for many realistic scenarios.
 
-$P^\pi_{i,j}=P^\pi(s_j|s_i)$ Probabilities to get to $s_j$ from $s_i$
-
-$r^\pi_i=R^\pi(s_i)$ Rewards of states
-
-$\mu_0=p(s_i)$ Initial state probabilities
-
-$\mu_\theta = \mu_0 + \gamma P^T \mu_\theta$
-
-$(I - \gamma P^T)\mu_\theta = \mu_0 \implies \mu_\theta^T = \mu_0(I-\gamma P)^{-1}$
-
-$J(\theta) = \mu_0^T \nabla_\theta V^\pi$ s.t. $V^\pi = r^\pi = \gamma P^\pi V^\pi$
-
-$\nabla_\theta J(\theta) = \mu_0^T \nabla_\theta V^\pi$
-
-$\nabla_\theta J(\theta) = \mu_0^T (I - \gamma P^\theta)^{-1} (\nabla_\theta r + \gamma (\nabla_\theta P^\theta) V^\theta) = \mu_\theta^T (\nabla_\theta r^\theta + \gamma (\nabla_\theta P^\theta) V^\theta)$
+\paragraph{Likelihood-ratio gradients.} 
+For this derivation, we will change the perspective a bit requiring some additional definitions. 
+We define ${\mu_\pi}_i = \sum_{t=0}^{\infty} p(s_t=i | s_0, \pi)$ \cite{puterman2014markov} as the state distribution, though it does not sum up to one without normalization.
+Note that $\mu_\pi$ is equivalent to the discounted state visit count $d^\pi$ introduced by Sutton et al.
+\cite{Sutton:1999:PGM:3009657.3009806}.
+We define $P_\pi$ as the transition matrix, i.e. ${P_\pi}_{i,j}=p(s_j|s_i,\pi)$.
+$r_\pi$ contains the mean rewards for all states. 
+$\mu_0$ is the initial state distribution.
+$\psi^\pi$ denotes some score function with $\psi_\pi(\tau) \propto \mathcal{R}^\tau$, see section \ref{sec:pge}, 
+From this, we derive
+\begin{align}
+	\mu_\pi &= \mu_0 + P_\pi^T \mu_\pi \nonumber \\
+	(I-P^T)\mu_\pi &= \mu_0 \nonumber \\
+	\mu_\pi^T &= \mu_u (I-P_\pi)^{-1} ,
+\end{align}
+so we can reformulate the problem as 
+\begin{align}
+	&J(\theta) = \mu_0^T \psi^\pi \\
+	\text{s.t.}\ &\psi^\pi = r^\pi = \gamma P^\pi \psi^\pi . \nonumber
+\end{align}
+From which it follows, that
+\begin{align*}
+	\nabla_\theta J(\theta) &= \mu_0^T \nabla_\theta \psi_\pi
+	\\ &= \mu_0^T (I - \gamma P_\pi)^{-1} (\nabla_\theta r + \gamma (\nabla_\theta P_\pi) \psi_\pi)
+	\\ &= \mu_\pi^T (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) \psi_\pi)
+	\\ &= \sum \mu_\theta(s) \nabla_\theta\pi(a|s) Q(s,a)
+	\\ &= \sum \mu_\theta(s) \pi(a|s) \nabla_\theta\log\pi(a|s) Q(s,a)
+\end{align*}
 
 $r_i^\theta = \int_\mathbb{A} r(s,a)\pi_\theta(a|s) da$
-
-$\nabla_\theta J(\theta) = \sum \mu_\theta(s) \nabla_\theta\pi(a|s) Q(s,a)$
-
-LR trick
-
-$ = \sum \mu_\theta(s) \pi(a|s) \nabla_\theta\log\pi(a|s) Q(s,a)$
 

### 2019-03-06 pge-mattis.tex
Update LR derivation (still in progress)

@@
 For discrete problems, we assume $\mathbb{S} \equiv \{0,1,2,\ldots\}$ and $\mathbb{A} \equiv \{0,1,2,\ldots\}$ without loss of generality.
-A trajectory $\tau_{\pi_\theta}^h := (s_0, a_0, $ $s_1, a_1, \dots, s_h, a_h)$ is generated by drawing $s_0 \sim p(x_0)$ according to the distribution over initial states $p(x_0)$, and successively sampling $a_t \sim \pi_\theta(a_t|s_t)$ according to the policy $\pi_\theta$, and $s_{t+1} \sim p(s_{t+1}|s_t,a_t)$ until the horizon $h$ is reached.
+A trajectory $\tau_{\pi_\theta}^h := (s_0, a_0, $ $s_1, a_1, \dots, s_h, a_h)$ is generated by drawing $s_0 \sim p(x_0)$ according to the distribution over initial states $p(s_0)$, and successively sampling $a_t \sim \pi_\theta(a_t|s_t)$ according to the policy $\pi_\theta$, and $s_{t+1} \sim p(s_{t+1}|s_t,a_t)$ until the horizon $h$ is reached.
 At each time step, we receive a reward according to $r_t = r(s_t, a_t, s_{t+1})$. 
@@
 
+\section{Policy Gradient Estimation}
+\label{sec:pge}
+
 \paragraph{Finite-difference gradients.} TODO
@@
 Since finite-difference gradients suffer from the curse of dimensionality, they are too inefficient for many realistic scenarios.
-
-\paragraph{Likelihood-ratio gradients.} 
-For this derivation, we will change the perspective a bit requiring some additional definitions. 
-We define ${\mu_\pi}_i = \sum_{t=0}^{\infty} p(s_t=i | s_0, \pi)$ \cite{puterman2014markov} as the state distribution, though it does not sum up to one without normalization.
-Note that $\mu_\pi$ is equivalent to the discounted state visit count $d^\pi$ introduced by Sutton et al.
-\cite{Sutton:1999:PGM:3009657.3009806}.
-We define $P_\pi$ as the transition matrix, i.e. ${P_\pi}_{i,j}=p(s_j|s_i,\pi)$.
-$r_\pi$ contains the mean rewards for all states. 
-$\mu_0$ is the initial state distribution.
-$\psi^\pi$ denotes some score function with $\psi_\pi(\tau) \propto \mathcal{R}^\tau$, see section \ref{sec:pge}, 
-From this, we derive
-\begin{align}
-	\mu_\pi &= \mu_0 + P_\pi^T \mu_\pi \nonumber \\
-	(I-P^T)\mu_\pi &= \mu_0 \nonumber \\
-	\mu_\pi^T &= \mu_u (I-P_\pi)^{-1} ,
-\end{align}
-so we can reformulate the problem as 
-\begin{align}
-	&J(\theta) = \mu_0^T \psi^\pi \\
-	\text{s.t.}\ &\psi^\pi = r^\pi = \gamma P^\pi \psi^\pi . \nonumber
-\end{align}
-From which it follows, that
-\begin{align*}
-	\nabla_\theta J(\theta) &= \mu_0^T \nabla_\theta \psi_\pi
-	\\ &= \mu_0^T (I - \gamma P_\pi)^{-1} (\nabla_\theta r + \gamma (\nabla_\theta P_\pi) \psi_\pi)
-	\\ &= \mu_\pi^T (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) \psi_\pi)
-	\\ &= \sum \mu_\theta(s) \nabla_\theta\pi(a|s) Q(s,a)
-	\\ &= \sum \mu_\theta(s) \pi(a|s) \nabla_\theta\log\pi(a|s) Q(s,a)
-\end{align*}
-
-$r_i^\theta = \int_\mathbb{A} r(s,a)\pi_\theta(a|s) da$
-
-Where we used $\nabla_\theta \pi_\theta(a|s) = \pi_\theta(a|s)\nabla_\theta\log\pi_\theta(a|s)$, obtained from the likelihood ratio, which gives us the likelihood-ratio gradient,
-\begin{equation}
-  \nabla_\theta \mathbb{E}_{\pi_\theta}\left[\psi(\tau)\right] = \mathbb{E}_{\pi_\theta}\left[\nabla_\theta{\log\pi_\theta(a|s)}\psi(\tau)\right].
-  \label{eqn:like-grad}
-\end{equation}
-intuitively meaning we should to increase the probability of actions judged as good by the score function $\psi$. We present some frequently used score functions in section \ref{sec:pge}.
-
-\section{Policy Gradient Estimation}
-\label{sec:pge}
-
-In this section, we discuss some approaches to calculate the estimated policy gradient $\nabla_\theta J(\theta) = \mathbb{E}_{\pi_\theta}\left[\nabla_\theta{\log\pi_\theta(a|s)}\psi(a|s)\right]$ using different score functions to substitute $\psi$.
-
-\paragraph{Immediate updates.} 
-The first class of algorithms developed to update a policy directly instead of deriving it from a value function approximation were REINFORCE \cite{Williams92simplestatistical} algorithms.
-REINFORCE uses immediate updates with a baseline $b$ in its gradient steps:
-\begin{equation}
-  \psi_t \leftarrow r_t-b.
-\end{equation}
-
-Obviously, this does not involve temporal information. 
-To credit temporal information $\psi_t$ needs to involve multiple time steps, meaning we need to delay updates until we have enough information to evaluate the gradient.
-
-\paragraph{Episodic updates.} 
-A common way of crediting temporal information in score function is to sample trajectories, also called roll-outs, and to updates using the complete roll-out. 
-We call these approaches episodic. 
-When we do this, we sum up the immediate updates, which we can rearrange to get
-\begin{equation}
-  \sum_{t=0}^h \nabla_\theta J(\theta) = \mathbb{E}_{\pi_\theta}\left[\sum_{t=0}^h \nabla_\theta{\log\pi_\theta(a|s)}\psi(\tau)\right].
-  \label{eqn:egrad}
-\end{equation}
-
-Episodic REINFORCE \cite{Williams92simplestatistical}, the first approach to this concept performs roll-outs and takes only the final return:
-\begin{equation}
-  \nabla_\theta J(\theta) \approx \frac{1}{h} \sum_{t=0}^h \nabla_\theta\log\pi_\theta(a|s) \mathcal{R}_t^h.
-\end{equation}
-This is sometimes also called Monte-Carlo gradient estimation. 
-However, taking this a step further, it should be possible to attribute the reward at each step to the respective action. 
-A very basic episodic score function that fits into our framework in \eqref{eqn:egrad} is to use the discounted rewards, similar to \eqref{eqn:acc-reward}:
-\begin{equation}
-  \psi_t \leftarrow \gamma^t r_t.
-\end{equation}
-If we only take batches of episodes starting at time $t'$, we can also use only the rewards following a chosen action $a_{t'}$ by calculating only $\sum_{t=t'}^h \nabla_\theta J(\theta)$ with $\psi_t \leftarrow \gamma^{t'} r_{t}$.
-As we will see, this concept is also used in value functions \eqref{eqn:v}, and \eqref{eqn:q}.
-
-Given $r_t > 0, \forall t=0,\dots,h$, these approaches can only increase action probabilities. 
-Obviously, we normalize to ensure $\forall s \in \mathbb{S}: \int_\mathbb{A}{\pi(a|s)da} = 1$. 
-This means actions can only become less probable in relation to other actions. 
-We find that this introduces a lot of variance when learning from samples \cite{Sutton:1999:PGM:3009657.3009806}. 
-One approach to counter the variance is to introduce a baseline
-\begin{equation}
-  \psi_t \leftarrow \gamma^{t} (r_{t} - b(s_t))
-\end{equation}
-The baseline can represent any prior knowledge or assumption about states, e.g., their mean value. 
-Though we find that there is a specific way to use this baseline which fits surprisingly well into this framework.
 
@@
 
-Approximating this function by an unbiased estimator $f_w(s_t, a_t) = \hat{Q}^\pi(s_t, a_t)$, we can also use this estimation as a score function $\psi \leftarrow \hat{Q}$.
-Sutton et al.
+\paragraph{Likelihood-ratio gradients.} 
+For this derivation, we will change the perspective a bit requiring some additional definitions. 
+We define ${\mu_\pi}_i = \sum_{t=0}^{\infty} p(s_t=s_i | s_0, \pi)$ \cite{puterman2014markov} as the state distribution, though it does not sum up to one without normalization.
+Note that $\mu_\pi$ is equivalent to the discounted state visit count $d^\pi$ introduced by Sutton et al.
+\cite{Sutton:1999:PGM:3009657.3009806}.
+We define $P_\pi$ as the transition matrix, i.e. ${P_\pi}_{i,j}=p(s_j|s_i,\pi)$.
+$r_\pi$ contains the mean rewards for all states given by ${r_\pi}_i = \int_\mathbb{A} r(s_i,a)\pi_\theta(a|s_i) da$. 
+$\mu_0$ is the initial state distribution.
+From which it follows that $V_\pi = \mu_\pi r_\pi$
+From this, we derive
+\begin{align}
+	\mu_\pi &= \mu_0 + P_\pi^T \mu_\pi \nonumber \\
+	(I-P^T)\mu_\pi &= \mu_0 \nonumber \\
+	\mu_\pi^T &= \mu_0^T (I-P_\pi)^{-1} ,
+\end{align}
+so we can reformulate the problem as 
+\begin{align}
+	\underset{\theta}{\text{max}}\ J(\theta) &= \mu_0^T r_\pi \\
+	\text{s.t.}\ r_\pi &= \gamma P_\pi V_\pi . \nonumber
+\end{align}
+From which we can derive 
+\begin{align*}
+	\nabla_\theta J(\theta) &= \nabla_\theta \mu_0^T r_\pi
+	\\ &= \mu_0^T \nabla_\theta \gamma P_\pi V_\pi
+	\\ &= \mu_0^T (I - \gamma P_\pi)^{-1} (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi)
+	\\ &= \mu_\pi^T (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) r_\pi)
+	\\ 
+	\\ &= \sum \mu_\theta(s) \nabla_\theta\pi(a|s) Q(s,a)
+	\\ &= \sum \mu_\theta(s) \pi(a|s) \nabla_\theta\log\pi(a|s) Q(s,a).
+\end{align*}
+Where we used $\nabla_\theta \pi_\theta(a|s) = \pi_\theta(a|s)\nabla_\theta\log\pi_\theta(a|s)$, obtained from the likelihood ratio, which gives us the likelihood-ratio gradient,
+\begin{equation}
+	\nabla_\theta J(\theta) = \mathbb{E}_{\pi_\theta}\left[\nabla_\theta{\log\pi_\theta(a|s)}Q_\pi(s,a)\right].
+	\label{eqn:like-grad}
+\end{equation}
+intuitively meaning we should to increase the probability of actions that lead to higher Q-values.
+
+Obviously, we do not have the true $Q_\pi(s,a)$, thus we need to approximate it by $\hat{Q}_\pi(s,a)$. 
+We draw $a \sim \pi(a|s)$, starting in state $s_0 \sim p(s_0)$ and match a function estimator to our observations.
+Approximating $Q_\pi(s,a)$ by an unbiased estimator $f_w(s_t, a_t) \equiv \hat{Q}_\pi(s_t, a_t)$, Sutton et al.
 \cite{Sutton:1999:PGM:3009657.3009806} show that using this function approximation we will converge to the same local optimum.
 However, they also find this requires $\int_\mathbb{S}{\pi(s,a)f_w(s,a)da} = 0$ for each state. 
-Given this assumption, the function estimator $f_w$ resembles an advantage function
+Given this assumption, the function estimator $f_w$ is limited to an advantage function
 \begin{equation}
@@
 \end{equation}
-as in \cite{1993b}.
+as in \cite{1993b}, which is mean-zero. We present the effect of this in greater detail in \ref{sec:ac}, but it does not critically affect the following findings.
+
+\paragraph{Immediate updates.} 
+The first class of algorithms developed to update a policy directly were REINFORCE \cite{Williams92simplestatistical} algorithms.
+REINFORCE uses immediate updates with a baseline $b$ in its gradient steps:
+\begin{equation}
+  \psi_t \leftarrow r_t-b.
+\end{equation}
+
+Obviously, this does not involve temporal information, and is very biased.
+To reduce the estimate's bias, we can do updates from completed roll-outs.
+
+\paragraph{Episodic updates.} 
+A common way of reducing the bias in value estimation is to sample trajectories, also called roll-outs, and to updates using the complete roll-out. 
+We call these approaches episodic. 
+When we do this, we sum up the immediate updates, which we can rearrange to get
+\begin{equation}
+  \sum_{t=0}^h \nabla_\theta J(\theta) = \mathbb{E}_{\pi_\theta}\left[\sum_{t=0}^h \nabla_\theta{\log\pi_\theta(a|s)}Q_\pi(s,a)\right].
+  \label{eqn:egrad}
+\end{equation}
+
+Episodic REINFORCE \cite{Williams92simplestatistical}, the first approach to this concept performs roll-outs and takes only the final return:
+\begin{equation}
+  \nabla_\theta J(\theta) \approx \frac{1}{h} \sum_{t=0}^h \nabla_\theta\log\pi_\theta(s) Q(s_t,a_t).
+\end{equation}
+This is sometimes also called Monte-Carlo gradient estimation. 
+However, taking this a step further, it should be possible to attribute the reward at each step to the respective action.
+
+Given $r_t > 0, \forall t=0,\dots,h$, these approaches can only increase action probabilities. 
+Obviously, we normalize to ensure $\forall s \in \mathbb{S}: \int_\mathbb{A}{\pi(a|s)da} = 1$. 
+This means actions can only become less probable in relation to other actions. 
+We find that this introduces a lot of variance when learning from samples \cite{Sutton:1999:PGM:3009657.3009806}. 
+One approach to counter the variance is to introduce a baseline
+\begin{equation}
+  \psi_t \leftarrow \gamma^{t} (r_{t} - b(s_t))
+\end{equation}
+The baseline can represent any prior knowledge or assumption about states, e.g., their mean value. 
+Though we find that there is a specific way to use this baseline which fits surprisingly well into this framework.
 

### 2019-03-07 pge-mattis.tex
Rewrite abstract and some refactoring

@@
 
-TODO
-
-Policy gradient methods use a policy model to decide which actions maximize the expected accumulated reward received in a given environment. 
-Policies can be deterministic but are often stochastic to accommodate for the nature of the given environment. 
-One class of problems that is frequently chosen to evaluate these methods are continuous control problems.
-Depending on the environment the formal design of an optimal controller can be infeasible or very labor-intensive.
-Policy gradient methods can provide useful control models while aiming to optimize their data efficiency, thus requiring less resources for the learning process. 
-The efficiency of these methods highly depends on the direction of change, i.e., the gradient they estimate. 
-The goal of policy gradient estimation is to estimate a function that predicts how the policy should be changed to solve the reinforcement learning problem of maximizing the accumulated reward of an agent.
-Thus, different functions for estimating the gradient have been proposed, some of which outperform others in specific areas. 
-This survey gives a general introduction to policy gradient, and gives an overview of existing gradient estimation approaches.
+The goal of policy gradient approaches is to find a policy in a given class of policies which maximizes the expected return.
+Given a differentiale model of the policy, e.g. exponential policies, we want to apply a gradient-ascent technique to reach the local optimum. 
+The main issue is that the policy gradient with respect to the expected return is not available, and we need to estimate it.
+Gradient ascent has strong guarantees on convergence, and is theoretically well researched.
+However, the main weakness of policy gradient algorithms is sample efficiency, as they tend to require on-policy data, and conservative learning rates.
+Thus, most research is focused on finding algorithms with improved sample efficiency.
+This paper provides a deep introduction to policy gradient that shows the development of policy gradient approaches, and should enable the reader to understand current research in the area.
 
@@
 
-Distributions often used as policies are Gibbs policies $\pi_\theta(a|s) = \frac{exp(\phi(s,a)^T\theta)}{\sum_b exp(\phi(s,b)^T\theta)}$ \cite{Sutton:1999:PGM:3009657.3009806,Bagnell2004LearningD} for discrete problems, and Gaussian policies $\pi_\theta(a|s) = N(\phi(s,a)^T\phi_1,\phi_2)$ for continuous problems, where $\phi_2$ is an exploration parameter \cite{Williams92simplestatistical,peter:article:1996}.
-$\phi(s,a)$ is the vector of basis functions on the state-action pair.
+The most frequently used policy classes in policy gradient approaches are Gibbs policies $\pi_\theta(a|s) = \frac{exp(\phi(s,a)^T\theta)}{\sum_b exp(\phi_1(s,b)^T\theta)}$ \cite{Sutton:1999:PGM:3009657.3009806,Bagnell2004LearningD} for discrete problems, and Gaussian policies $\pi_\theta(a|s) = \mathcal{N}(\phi(s,a)^T\theta_1,\theta_2)$ for continuous problems, where $\theta_2$ is an exploration parameter \cite{Williams92simplestatistical,peter:article:1996}, and $\phi(s,a)$ is the vector of basis functions on the state-action pair.
 
@@
 
-\paragraph{Finite-difference gradients.} TODO
-
-Since finite-difference gradients suffer from the curse of dimensionality, they are too inefficient for many realistic scenarios.
+\paragraph{Finite-difference gradients.} 
+A simple approach for gradient estimation is to choose a small $\delta\theta$, and evaluate the new policy given the slightly changed parameters as in 
+\begin{equation}
+	\nabla_\theta J(\theta) \approx \frac{J(\theta+\delta\theta)J(\theta-\delta\theta)}{2\delta\theta}.
+\end{equation} 
+This can lend a good estimate of the gradient given a small $\delta$, and is generally called the symmetric derivative. 
+However, finite-difference gradients suffer from the curse of dimensionality, and can require very small $\delta\theta$, thus they only work well in specific scenarios, but should not be discarded due to simplicity.
 
@@
 
-In general, we need to approximate this function. 
-Obviously, REINFORCE does this approximation by a simple Monte-Carlo sample. 
-However, there is a formulation of policy gradient methods that implements this approximation problem explicitly called actor-critic, which we discuss in \ref{sec:ac}.
-
-Another frequently used function is the state-action value function. 
-Instead of the expected accumulated reward starting from state $s_t$, this function gives the expected accumulated reward given an action $a_t$ is selected
+As we will see, this function also gives us the true gradient of $J(\theta$, though in general we need to estimate it. 
+In addition to the value function, we also define the state-action value function, often called Q-function. 
+Instead of the expected accumulated reward starting from state $s_t$, this function gives the expected accumulated reward given an action $a_t$ is selected in state $s_t$, 
 \begin{equation}
@@
 \end{equation}
+
+Using the value function, and the Q-function, we can derive a better estimate of the policy gradient.
 
@@
 \begin{align}
-	\underset{\theta}{\text{max}}\ J(\theta) &= \mu_0^T r_\pi \\
-	\text{s.t.}\ r_\pi &= \gamma P_\pi V_\pi . \nonumber
+	\underset{\theta}{\text{max}}\ J(\theta) &= \mu_0^T V_\pi \\
+	\text{s.t.}\ V_\pi &= r_\pi + \gamma P_\pi V_\pi . \nonumber
 \end{align}
-From which we can derive 
+Where $V_\pi = \left[ V_\pi(s_0), V_\pi(s_1), \ldots \right]^T$. We derive 
 \begin{align*}
-	\nabla_\theta J(\theta) &= \nabla_\theta \mu_0^T r_\pi
+	\nabla_\theta J(\theta) &= \nabla_\theta \mu_0^T V_\pi
 	\\ &= \mu_0^T \nabla_\theta \gamma P_\pi V_\pi
@@
 	\\ &= \mu_\pi^T (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) r_\pi)
-	\\ 
-	\\ &= \sum \mu_\theta(s) \nabla_\theta\pi(a|s) Q(s,a)
-	\\ &= \sum \mu_\theta(s) \pi(a|s) \nabla_\theta\log\pi(a|s) Q(s,a).
+	\\ &\equiv \sum\nolimits_i d^\pi(s_i) \nabla_\theta\pi(a|s_i) Q(s_i,a)
+	\\ &= \sum\nolimits_i d^\pi(s_i) \pi(a|s_i) \nabla_\theta\log\pi(a|s_i) Q_\pi(s_i,a).
 \end{align*}
@@
 \begin{equation}
-	\nabla_\theta J(\theta) = \mathbb{E}_{\pi_\theta}\left[\nabla_\theta{\log\pi_\theta(a|s)}Q_\pi(s,a)\right].
+	\nabla_\theta J(\theta) = \mathbb{E}_{\pi}\left[\nabla_\theta{\log\pi_\theta(a|s)}Q_\pi(s,a)\right].
 	\label{eqn:like-grad}

### 2019-03-07 pge-mattis.tex
Fix some misunderstandings

@@
 We define states $s \in \mathbb{S}$, actions $a \in \mathbb{A}$, and rewards $r \in \mathbb{R}$. 
-For discrete problems, we assume $\mathbb{S} \equiv \{0,1,2,\ldots\}$ and $\mathbb{A} \equiv \{0,1,2,\ldots\}$ without loss of generality.
-A trajectory $\tau_{\pi_\theta}^h := (s_0, a_0, $ $s_1, a_1, \dots, s_h, a_h)$ is generated by drawing $s_0 \sim p(x_0)$ according to the distribution over initial states $p(s_0)$, and successively sampling $a_t \sim \pi_\theta(a_t|s_t)$ according to the policy $\pi_\theta$, and $s_{t+1} \sim p(s_{t+1}|s_t,a_t)$ until the horizon $h$ is reached.
-At each time step, we receive a reward according to $r_t = r(s_t, a_t, s_{t+1})$. 
-A trajectory can also be called \textit{roll-out} or \textit{episode}, though the term episode implies it ends in a terminal state.
-
+A trajectory $\tau := (s_0, a_0, $ $s_1, a_1, \dots, s_h, a_h)$ is generated by drawing $s_0 \sim \mu_0(s_0)$ according to the distribution over initial states $\mu_0(s_0)$, and successively sampling $a_t \sim \pi_\theta(a_t|s_t)$ according to the policy $\pi_\theta$, and $s_{t+1} \sim p(s_{t+1}|s_t,a_t)$ until the horizon $h$, or a terminal state is reached.
+At each time step, we receive a reward according to $r_t = r(s_t, a_t) \equiv \mathbb{E}_{s'}\left[r(s_t,a_t,s')\right]$. 
+A trajectory can also be called roll-out or episode, though the term episode implies it ends in a terminal state.
 We assume a Markov Decision Process (MDP), meaning the probability distribution of the next states is independent of past states $s_{0:t-1}$ given the present state $s_t$ and action $a_t$, 
@@
 \end{equation}
-Where we define $i:j$ with $i,j \in \mathbb{N}, i < j$ as an index over all integers from $i$ to $j$, i.e., $s_{i:j} := s_i, s_{i+1}, \dots, s_j$.
+Where we define $i:j$ with $i,j \in \mathbb{N}, i < j$ as an index over all integers from $i$ to $j$, i.e., $s_{i:j} \equiv s_i, s_{i+1}, \dots, s_j$. 
+We assume no additional prior knowledge about the environment, thus we assume the probability of a trajectory is 
+\begin{equation}
+	p(\tau) = \mu_0(s_0) \prod_{t=0}^{h-1} p(s_{t+1}|s_{t}, a_{t}) \pi(a_t|s_{t}).
+\end{equation}
 
@@
 Thus, we search $\nabla_\theta J(\theta) := \nabla_\theta \mathbb{E}_{p(\tau)}\left[\mathcal{R}^\tau\right]$, to make a policy gradient step according to $\theta_{k+1} = \theta_k + \alpha_k \nabla_\theta J(\theta)$, where $\alpha_k$ denotes a learning rate.
-It is shown that for $\lim_{k\to\infty}\alpha_k = 0$, and $\sum_{k=0}^\infty \alpha_k$ we are guaranteed to converge to a local optimum \cite{Sutton:1999:PGM:3009657.3009806}.
 
@@
 \cite{Sutton:1999:PGM:3009657.3009806}.
-We define $P_\pi$ as the transition matrix, i.e. ${P_\pi}_{i,j}=p(s_j|s_i,\pi)$.
-$r_\pi$ contains the mean rewards for all states given by ${r_\pi}_i = \int_\mathbb{A} r(s_i,a)\pi_\theta(a|s_i) da$. 
-$\mu_0$ is the initial state distribution.
+Further, we define $P_\pi$ as the transition matrix, i.e.
+${P_\pi}_{i,j}=p(s_j|s_i,\pi)$, $r_\pi$ as the mean rewards for all states given by ${r_\pi}_i = \int_\mathbb{A} r(s_i,a)\pi_\theta(a|s_i) da$, and $\mu_0 = \left[\mu_0(s_0),\mu_0(s_1),\ldots\right]^T$ as a vector representing the initial state distribution.
 From which it follows that $V_\pi = \mu_\pi r_\pi$
@@
 We draw $a \sim \pi(a|s)$, starting in state $s_0 \sim p(s_0)$ and match a function estimator to our observations.
+It is shown that for $\lim_{k\to\infty}\alpha_k = 0$, and $\sum_{k=0}^\infty \alpha_k$ we are guaranteed to converge to a local optimum \cite{Sutton:1999:PGM:3009657.3009806}.
 Approximating $Q_\pi(s,a)$ by an unbiased estimator $f_w(s_t, a_t) \equiv \hat{Q}_\pi(s_t, a_t)$, Sutton et al.
-\cite{Sutton:1999:PGM:3009657.3009806} show that using this function approximation we will converge to the same local optimum.
-However, they also find this requires $\int_\mathbb{S}{\pi(s,a)f_w(s,a)da} = 0$ for each state. 
-Given this assumption, the function estimator $f_w$ is limited to an advantage function
-\begin{equation}
-	A^{\pi}(s_t, a_t) := Q^{\pi}(s_t, a_t) - V^{\pi}(s_t)
-	\label{eqn:adv}
-\end{equation}
-as in \cite{1993b}, which is mean-zero. We present the effect of this in greater detail in \ref{sec:ac}, but it does not critically affect the following findings.
+\cite{Sutton:1999:PGM:3009657.3009806} show that using this function approximation we will converge to the real local optimum.
 
 \paragraph{Immediate updates.} 
-The first class of algorithms developed to update a policy directly were REINFORCE \cite{Williams92simplestatistical} algorithms.
-REINFORCE uses immediate updates with a baseline $b$ in its gradient steps:
+The first class of algorithms developed to update a policy directly are called REINFORCE \cite{Williams92simplestatistical}.
+REINFORCE collects a complete episode, at which point we can calculate the actual state-action value by traversing backwards over the trajectory, and estimates:
 \begin{equation}
-  \psi_t \leftarrow r_t-b.
+	\nabla_\theta J(\theta) \approx \nabla_\theta\log\pi_\theta(s) \left(Q(s_t,a_t) - b \right),
+	\label{eqn:reinforce}
+\end{equation}
+where $b$ is some baseline. 
+Peters \& Schaal \cite{4867} find that an estimate of the optimal baseline can be calculated by
+\begin{equation}
+	b = \frac
+		{\left\langle 
+			\left(\sum_{t=0}^h \nabla_\theta \log\pi_\theta(a_t|s_t) \right)^2 \sum_{l=0}^h a_l r_l 
+		\right\rangle}
+		{\left\langle
+			\left(\sum_{t=0}^h \nabla_\theta \log\pi_\theta(a_t|s_t) \right)^2
+		\right\rangle}
 \end{equation}
 
-Obviously, this does not involve temporal information, and is very biased.
-To reduce the estimate's bias, we can do updates from completed roll-outs.
+However, this estimate is very biased.
+To reduce the estimate's bias, we can compute updates from a complete episode.
 
@@
 
-Episodic REINFORCE \cite{Williams92simplestatistical}, the first approach to this concept performs roll-outs and takes only the final return:
+Episodic REINFORCE \cite{Williams92simplestatistical}, the first approach to this concept, performs roll-outs and takes the final return, at which point we can calculate the actual state-action value by traversing backwards over the trajectory, and estimates
 \begin{equation}
@@
 This is sometimes also called Monte-Carlo gradient estimation. 
-However, taking this a step further, it should be possible to attribute the reward at each step to the respective action.
-
 Given $r_t > 0, \forall t=0,\dots,h$, these approaches can only increase action probabilities. 
@@
 This means actions can only become less probable in relation to other actions. 
-We find that this introduces a lot of variance when learning from samples \cite{Sutton:1999:PGM:3009657.3009806}. 
+We find that this introduces more variance when learning from samples \cite{Sutton:1999:PGM:3009657.3009806}. 
 One approach to counter the variance is to introduce a baseline
 \begin{equation}
-  \psi_t \leftarrow \gamma^{t} (r_{t} - b(s_t))
+  \hat{Q}(s,a) \approx \gamma^{t} (r_{t} - b(s_t)).
 \end{equation}
-The baseline can represent any prior knowledge or assumption about states, e.g., their mean value. 
-Though we find that there is a specific way to use this baseline which fits surprisingly well into this framework.
+The baseline can represent any prior knowledge or assumption about states. 
+A common way to use this baseline is to use the value function, which gives us the so-called advantage function
+
+\begin{equation}
+	A_{\pi}(s_t, a_t) := Q_{\pi}(s_t, a_t) - V_{\pi}(s_t).
+	\label{eqn:adv}
+\end{equation}
 
@@
 This means, the actor consists only of the policy, while the critic is focused on estimating a score for the actions taken. 
-By that concept, observations of $s_t$ from the environment are given to the actor only to decide the next action, and to the critic only to improve its function estimation with the respective $r_t$. 
+By that concept, observations of the environment are given to the actor only to decide the next action, and to the critic only to improve its function estimation with the respective rewards. 
 Figure \ref{fig:ac} shows the general structure of an actor-critic algorithm.
 
-\paragraph{Actor.} 
-TODO
+The critic estimates a state-action value function as defined in \eqref{eqn:q}. 
+\cite{Sutton:1999:PGM:3009657.3009806}, and \cite{Konda:2003:AA:942271.942292} find that the estimation $f_w^\pi(s,a) \approx Q_\pi(s,a)$ does not affect the unbiasedness of the gradient estimate under some restrictions.
+Specifically, this holds for $f_w^\pi(s,a) = {\nabla_\theta \log\pi(a|s)}^T w$, thus $f_w^\pi(s,a)$ being a linear function parameterized by the vector $w$.
+This guarantees that the function estimator does not cause divergence, and really enables recent research in reinforcement learning for continuous control problems, e.g., in humanoid robotics.
 
-\paragraph{Critic.} 
-The critic estimates a value function as defined in \eqref{eqn:v}. 
-\cite{Sutton:1999:PGM:3009657.3009806}, and \cite{Konda:2003:AA:942271.942292} find that the estimation $f_w^\pi(s,a) \approx Q^\pi(s,a)$ does not affect the unbiasedness of the gradient estimate under some restrictions.
-Specifically, this holds for $f_w^\pi(s,a) = {\nabla_\theta \log\pi(a|s)}^T w$, thus $f_w^\pi(s,a)$ being a linear function parameterized by the vector $w$. 
-However, as we have seen in the last section, \ref{sec:pge}, this function estimator will have a mean of zero $\int_\mathbb{A} \nabla_\theta \pi(a|s)du $, thus it estimates an advantage function \ref{eqn:adv}.
-However, to estimate the advantage function, we require $V^\pi(s)$. 
-We will show how this leads to a new problem but still, this insight guarantees that function approximation does not cause divergence, and really enables recent research in reinforcement learning for continuous control problems, e.g., in humanoid robotics.
-
-Traditionally, the improvement is often done by Monte-Carlo sampling as in REINFORCE, or using temporal difference (TD) \cite{Sutton1988}, i.e., we use the temporal difference between the critic's estimations
+Traditionally, the improvement is often done by Monte-Carlo sampling as in REINFORCE \eqref{eqn:reinforce}, or using temporal difference (TD) \cite{Sutton1988}, i.e., we use the temporal difference between the critic's estimations
 \begin{equation}
-  r_t + \gamma \hat{V}^\pi(s_{t+1}) - \hat{V}^\pi(s_t).
+  r_t + \gamma \hat{V}_\pi(s_{t+1}) - \hat{V}_\pi(s_t).
 \end{equation}
 
-This is where we need to come back to the insight that from $f_w^\pi(s,a) = {\nabla_\theta \log\pi(a|s)}^T w$, it follows that $f_w^\pi(s,a)$ estimates the advantage function, so we are bootstrapping for $V^\pi(s,a)$.
-If we use temporal difference in this context, we run into a problem, as \eqref{eqn:adv} subtracts $\hat{V}^\pi(s_t)$, meaning we would only learn immediate rewards \cite{Peters_IICHR_2003}. 
+However, Sutton et al. \cite{1993b} find that this is only guaranteed to be unbiased, if $\int_\mathbb{S}{\pi(s,a)f_w(s,a)da} = 0$ for each state. 
+Given this assumption, the function estimator $f_w$ is limited to an advantage function \eqref{eqn:adv}, which requires bootstrapping for $V_\pi(s,a)$. 
+If we use temporal difference in this context, we run into a problem, as \eqref{eqn:adv} subtracts $\hat{V}_\pi(s_t)$, meaning we would only learn immediate rewards \cite{Peters_IICHR_2003}. 
 This would render the process biased. 
 Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806} and Konda et al. \cite{NIPS1999_1786} suggest estimating an action value function as in \eqref{eqn:q}. 
-We can approximate this $f_w^\pi$ by least-squares optimization over multiple $\hat{Q}^\pi(s,a)$ obtained from roll-outs. 
+We can approximate this $f_w^\pi$ by least-squares optimization over multiple $\hat{Q}_\pi(s,a)$ obtained from roll-outs. 
 However, Peters et al. \cite{4863} find that this approximation is highly reliant on the distribution of the training data. 
@@
 One can compare this to approximating a parabola by a line, whereby the approximation changes wildly depending on which part of the parabola is in the training data. 
-An approach to solve this bootstrapping problem is to rewrite the Bellman Equation using \eqref{eqn:adv} and \eqref{eqn:q} with $A^\pi(s,a) = f_w^\pi(s,a)$, $V^\pi(s) = \phi(s)^T v$, and a zero-mean error term $\epsilon(s_t,a_t,s_{t+1})$, we get
+An approach to solve this bootstrapping problem is to rewrite the Bellman Equation using \eqref{eqn:adv} and \eqref{eqn:q} with $A_\pi(s,a) = f_w^\pi(s,a)$, $V_\pi(s) = \phi(s)^T v$, and a zero-mean error term $\epsilon(s_t,a_t,s_{t+1})$, we get
 \begin{align}
-  Q^\pi(s,a) = A^\pi(s,a) + V^\pi(s) = r(s,a) + \gamma \int_\mathbb{S} p(s'|s,a)V^\pi(s')ds' \\
-  \nabla_\theta \log \pi(a_t|s_t)^T w + \phi(s_t)^T v = r(s_t,a_t) + \gamma \phi(s_{t+1})^T v + \epsilon(s_t,a_t,s_{t+1})
+  Q_\pi(s,a) = A_\pi(s,a) + V_\pi(s) &= r(s,a) + \gamma \int_\mathbb{S} p(s'|s,a)V_\pi(s')ds' \\
+  \nabla_\theta \log \pi(a_t|s_t)^T w + \phi(s_t)^T v &= r(s_t,a_t) + \gamma \phi(s_{t+1})^T v + \epsilon(s_t,a_t,s_{t+1})
 \end{align}
@@
 \begin{equation}
-  F_\theta = \int_\mathbb{S} p(s|\theta) \int_\mathbb{A} \pi_\theta(a|s) \nabla_\theta \log{\pi_\theta(a|s)} \nabla_\theta \log{\pi_\theta(a|s)}^T dads.
+  F_\theta = \int_\mathbb{S} d^\pi(s) \int_\mathbb{A} \pi_\theta(a|s) \nabla_\theta \log{\pi_\theta(a|s)} \nabla_\theta \log{\pi_\theta(a|s)}^T dads.
   \label{eqn:F}
 \end{equation}
-When we recall the definition \eqref{eqn:like-grad} of likelihood-ratio gradients, we see that setting $\psi$ as $\nabla_\theta \log \pi(a|s)^T w$, we get
+When we recall the definition \eqref{eqn:like-grad} of likelihood-ratio gradients, we see that setting $\hat{Q}(s,a) := \nabla_\theta \log \pi(a|s)^T w$, we get
 \begin{equation}
-  \nabla_\theta J(\theta) = \int_\mathbb{S} {p(s|\theta) \int_\mathbb{A} \pi_\theta(a|s) \nabla_\theta \log \pi(a|s) {\nabla_\theta \log \pi(a|s)}^T dads}\ w := F_\theta w.
+  \nabla_\theta J(\theta) = \int_\mathbb{S} {p(s|\theta) \int_\mathbb{A} \pi_\theta(a|s) \nabla_\theta \log \pi(a|s) {\nabla_\theta \log \pi(a|s)}^T dads}\ w \equiv F_\theta w.
   \label{eqn:J-equals-F}
@@
 
-Note that this approach abuses some specific conveniences about the problem setup to find an elegant solution.
-However, this is not the only way to approach this problem. 
-In this paper, we focus on one presentation to present the core idea, but the field of policy gradient optimization is very resourceful. 
-TODO 
+However, the natural gradient is similarly used by TODO: reference other algos using natural gradient and remove them from conclusion (TRPO?)
+
 In the conclusion \ref{sec:outro}, we attempt to show the versatility and relevance in current research, and give some directions for further reading.

### 2019-03-07 pge-mattis.tex
Rewrite conclusion

@@
 \end{equation}
-When we recall the definition \eqref{eqn:like-grad} of likelihood-ratio gradients, we see that setting $\hat{Q}(s,a) := \nabla_\theta \log \pi(a|s)^T w$, we get
+Using a value function estimator and calculating the natural gradient, we get the natural policy gradient algorithm (NPG) \cite{NIPS2017_7233}.
+But, if we recall the definition \eqref{eqn:like-grad} of likelihood-ratio gradients, we see that setting $\hat{Q}(s,a) := \nabla_\theta \log \pi(a|s)^T w$, we get
 \begin{equation}
@@
 \end{equation}
-
 From \eqref{eqn:nat-grad}, \eqref{eqn:F}, and \eqref{eqn:J-equals-F}, it follows that
@@
 Peters et al.
-\cite{4863} present this idea and suggest LSTD($\lambda$)-Q, a version of least-squares temporal difference learning  \cite{Boyan:1999:LTD:645528.657618}, as well as episodic natural actor-critic.
-
-However, the natural gradient is similarly used by TODO: reference other algos using natural gradient and remove them from conclusion (TRPO?)
-
-In the conclusion \ref{sec:outro}, we attempt to show the versatility and relevance in current research, and give some directions for further reading.
+\cite{4863} present this idea and suggest LSTD($\lambda$)-Q, a version of least-squares temporal difference learning  \cite{Boyan:1999:LTD:645528.657618}, as well as episodic natural actor-critic (eNAC).
 
@@
 
-TODO less references and stuff...
-
 In this paper, we have introduced policy gradient methods as a class of reinforcement learning algorithms. 
 We show why policy gradient methods are effective in these environments, and we give some intuitions about the concept. 
-Further, we show the core elements of policy gradient methods, and we discuss some intricacies that they bring, and some research development in the attempts of improving the efficiency and stability of policy gradients.
+Further, we show the core elements of policy gradient methods, discuss some intricacies the estimation of the policy gradient brings, and follow the research development in the attempts of improving the efficiency and stability of policy gradients.
+We show that we can reuse value-estimation approaches in actor-critic settings to improve gradient estimate through better policy evaluation.
+This lead to the introduction of the natural gradient as a way to iterate through policy space instead of parameter space, which improves sample efficiency, especially when the gradient in parameter space is very small, i.e., if there is a plateau.
 
-There are many possibilities for improvement as it can be seen as a very modularized and flexible. 
-We find that the theoretical basis for policy gradient focused approaches reaches far back with some fundamental algorithms like Levenberg-Marquardt \cite{levenbergmarquardt1978}, and the first presentation of REINFORCE \cite{Williams92simplestatistical}.
-Policy gradient methods were shown to be increasingly potent by authors such as Sutton \cite{Sutton:1999:PGM:3009657.3009806}, and still provide space for a lot of new work being published. 
-Algorithms such as the aforementioned Natural Actor-Critic \cite{4863} have been showing some of their potentials, but even so potential areas of improvement. 
-Schulman et al. \cite{DBLP:journals/corr/SchulmanMLJA15} present a form of advantage estimation using $\gamma$ as a parameter for bias-variance trade-off. 
-Fellows et al. \cite{fellows:icml18} use Fourier analysis to analytically improve policy gradients. 
-Abdolmaleki et al. \cite{abdolmaleki2018maximum} introduce MPO, an off-policy algorithm using expectation maximization, related to the approach of TRPO \cite{pmlr-v37-schulman15}. 
-Yet a different approach is given by Zhang et al.
-\cite{DBLP:journals/corr/abs-1808-03030}, who formulate policy optimization as a Wasserstein gradient flow problem, which moves into a Wasserstein manifold in the distribution space.
-This is comparable to approaches by Abdolmaleki et al. \cite{NIPS2015_5672}, which introduces a bound on the Kullback-Leibler divergence during policy search. 
 From the developments in recent research, it is fair to say that policy gradient methods play a major role in reinforcement learning.

### 2019-03-07 pge-mattis.tex
Minor fixes and ready for review

@@
 The goal of policy gradient approaches is to find a policy in a given class of policies which maximizes the expected return.
-Given a differentiale model of the policy, e.g. exponential policies, we want to apply a gradient-ascent technique to reach the local optimum. 
-The main issue is that the policy gradient with respect to the expected return is not available, and we need to estimate it.
+Given a differentiable model of the policy, we want to apply a gradient-ascent technique to reach the local optimum. 
 Gradient ascent has strong guarantees on convergence, and is theoretically well researched.
-However, the main weakness of policy gradient algorithms is sample efficiency, as they tend to require on-policy data, and conservative learning rates.
+The main issue is that the policy gradient with respect to the expected return is not available, thus we need to estimate it.
+For this reason, the biggest weakness of policy gradient algorithms is sample efficiency, as they tend to require on-policy data for the gradient estimate, and conservative learning rates for the convergence guarantees.
 Thus, most research is focused on finding algorithms with improved sample efficiency.
@@
 Also, they frequently require on-policy training and often converge slower, but can guaranteed find a local optimum given some constraints on the hyperparameters. 
-Options to model a policy gradient method are not only given by the policy model itself, but also in the way a policy improvement is computed \cite{DBLP:journals/corr/SchulmanMLJA15}. 
+Options to model a policy gradient method are not only given by the policy model itself, but also in the way a policy improvement is computed. 
 In this case, improvement means a step in the parameter space such that the policy under the new parameters will on average perform better than the old policy, i.e. improve its expected return. 
@@
 In section \ref{sec:pge}, we discuss different approaches to estimate the policy gradient. 
-Using our insights from section \ref{sec:pge}, we derive the actor-critic framework in section \ref{sec:ac}, which harnesses some value-based approaches for improved gradient updates. 
+Using our insights from section \ref{sec:pge}, we derive the actor-critic framework in section \ref{sec:ac}, which harnesses value-function estimation for improved gradient updates. 
 Then, in section \ref{sec:natural}, we introduce some gradient ascent methods that build on the approaches given in \ref{sec:pge}, refining the gradient estimation by Fisher's information matrix to get the natural gradient.
@@
 We define states $s \in \mathbb{S}$, actions $a \in \mathbb{A}$, and rewards $r \in \mathbb{R}$. 
-A trajectory $\tau := (s_0, a_0, $ $s_1, a_1, \dots, s_h, a_h)$ is generated by drawing $s_0 \sim \mu_0(s_0)$ according to the distribution over initial states $\mu_0(s_0)$, and successively sampling $a_t \sim \pi_\theta(a_t|s_t)$ according to the policy $\pi_\theta$, and $s_{t+1} \sim p(s_{t+1}|s_t,a_t)$ until the horizon $h$, or a terminal state is reached.
+A trajectory $\tau := (s_0, a_0, $ $s_1, a_1, \dots, s_T, a_T)$ is generated by drawing $s_0 \sim \mu_0(s_0)$ according to the distribution over initial states $\mu_0(s_0)$, and successively sampling $a_t \sim \pi_\theta(a_t|s_t)$ according to the policy $\pi_\theta$, and $s_{t+1} \sim p(s_{t+1}|s_t,a_t)$ until the horizon $T$, or a terminal state is reached.
 At each time step, we receive a reward according to $r_t = r(s_t, a_t) \equiv \mathbb{E}_{s'}\left[r(s_t,a_t,s')\right]$. 
@@
 \begin{equation}
-	p(\tau) = \mu_0(s_0) \prod_{t=0}^{h-1} p(s_{t+1}|s_{t}, a_{t}) \pi(a_t|s_{t}).
+	p_\pi(\tau) = \mu_0(s_0) \prod_{t=0}^{T-1} p(s_{t+1}|s_t, a_t) \pi(a_t|s_t).
 \end{equation}
@@
 Our goal with respect to episodes is to maximize the expectation of the total reward, also called expected return. 
-The total reward in the horizon $h$ is $\sum_{t=0}^{h} r_{t}$. 
+The total reward in the horizon $T$ is $\sum_{t=0}^{T} r_{t}$. 
 We introduce a discount factor $\gamma \in [0,1)$ to trade-off bias ($\gamma\to0$) versus variance ($\gamma\to1$) with a normalization factor of $1-\gamma$, which is often omitted though strongly recommended.
@@
 \begin{equation}
-  \mathcal{R}_0^h = \mathcal{R}^\tau := (1-\gamma) \sum_{t=0}^{h} \gamma^t r_t.
+  \mathcal{R}^\tau \equiv \mathcal{R}_0^T := (1-\gamma) \sum_{t=0}^{T} \gamma^t r_t.
   \label{eqn:acc-reward}
@@
 Since we have only limited knowledge of the performance of the policy, we need to approximate an optimal policy by estimating a gradient. 
-Thus, we search $\nabla_\theta J(\theta) := \nabla_\theta \mathbb{E}_{p(\tau)}\left[\mathcal{R}^\tau\right]$, to make a policy gradient step according to $\theta_{k+1} = \theta_k + \alpha_k \nabla_\theta J(\theta)$, where $\alpha_k$ denotes a learning rate.
+Thus, we search $\nabla_\theta J(\theta) := \nabla_\theta \mathbb{E}_{p_\pi(\tau)}\left[\mathcal{R}^\tau\right]$, to make a policy gradient step according to $\theta_{k+1} = \theta_k + \alpha_k \nabla_\theta J(\theta)$, where $\alpha_k$ denotes a learning rate.
 
@@
 This can lend a good estimate of the gradient given a small $\delta$, and is generally called the symmetric derivative. 
-However, finite-difference gradients suffer from the curse of dimensionality, and can require very small $\delta\theta$, thus they only work well in specific scenarios, but should not be discarded due to simplicity.
+However, finite-difference gradients suffer from the curse of dimensionality, and can require very small $\delta\theta$.
+Thus, they only work well in specific scenarios, but should not be discarded due to simplicity.
 
 \paragraph{Value functions.} 
-Given we know the actual value, i.e. the expected accumulated reward we will get when we start in a state $s_t$, this function can be used to evaluate the performance of our policy.
+Given we know the actual value, i.e. the expected return we will get starting from state $s_t$, this function can be used to evaluate the performance of our policy, and can be written as
 \begin{equation}
-	V^{\pi}(s_t) := \mathbb{E}_{\substack{s_{t+1:h} \\ a_{t:h}}}\left[\mathcal{R}_t^h\right].
+	V^{\pi}(s_t) := \mathbb{E}_{\substack{s_{t+1:h} \\ a_{t:h}}}\left[\mathcal{R}_t^T\right].
 	\label{eqn:v}
@@
 
-As we will see, this function also gives us the true gradient of $J(\theta$, though in general we need to estimate it. 
+As we will see, this function also gives us the true gradient of $J(\theta)$, though in general we need to estimate it. 
 In addition to the value function, we also define the state-action value function, often called Q-function. 
@@
 \begin{equation}
-	Q^{\pi}(s_t, a_t) := \mathbb{E}_{\substack{s_{t+1:h} \\ a_{t+1:h}}}\left[\mathcal{R}_t^h\right].
+	Q^{\pi}(s_t, a_t) := \mathbb{E}_{\substack{s_{t+1:h} \\ a_{t+1:h}}}\left[\mathcal{R}_t^T\right].
 	\label{eqn:q}
@@
 \paragraph{Likelihood-ratio gradients.} 
-For this derivation, we will change the perspective a bit requiring some additional definitions. 
+For this derivation, we will change the perspective a bit, requiring some additional definitions. 
 We define ${\mu_\pi}_i = \sum_{t=0}^{\infty} p(s_t=s_i | s_0, \pi)$ \cite{puterman2014markov} as the state distribution, though it does not sum up to one without normalization.
@@
 	\\ &= \mu_\pi^T (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) r_\pi)
-	\\ &\equiv \sum\nolimits_i d^\pi(s_i) \nabla_\theta\pi(a|s_i) Q(s_i,a)
-	\\ &= \sum\nolimits_i d^\pi(s_i) \pi(a|s_i) \nabla_\theta\log\pi(a|s_i) Q_\pi(s_i,a).
+	\\ &\equiv \sum\nolimits_{i,j} d^\pi(s_i) \nabla_\theta\pi(a_j|s_i) Q(s_i,a_j)
+	\\ &= \sum\nolimits_{i,j} d^\pi(s_i) \pi(a_j|s_i) \nabla_\theta\log\pi(a_j|s_i) Q_\pi(s_i,a_j).
 \end{align*}
@@
 \begin{equation}
-	\nabla_\theta J(\theta) = \mathbb{E}_{\pi}\left[\nabla_\theta{\log\pi_\theta(a|s)}Q_\pi(s,a)\right].
+	\nabla_\theta J(\theta) = \mathbb{E}_{\substack{\ s \sim d^\pi \\a \sim \pi}} \Big[\nabla_\theta{\log\pi_\theta(a|s)}Q_\pi(s,a)\Big].
 	\label{eqn:like-grad}
@@
 \begin{equation}
-	\nabla_\theta J(\theta) \approx \nabla_\theta\log\pi_\theta(s) \left(Q(s_t,a_t) - b \right),
+	\nabla_\theta J(\theta) \approx \nabla_\theta\log\pi_\theta(s) \left(Q(s_t,a_t) - b_\tau \right).
 	\label{eqn:reinforce}
 \end{equation}
-where $b$ is some baseline. 
+where $b_\tau$ is some baseline. 
 Peters \& Schaal \cite{4867} find that an estimate of the optimal baseline can be calculated by
 \begin{equation}
-	b = \frac
+	b_\tau = \frac
 		{\left\langle 
-			\left(\sum_{t=0}^h \nabla_\theta \log\pi_\theta(a_t|s_t) \right)^2 \sum_{l=0}^h a_l r_l 
+			\left(\sum_{t=0}^T \nabla_\theta \log\pi_\theta(a_t|s_t) \right)^2 \sum_{t'=0}^T a_{t'} r_{t'} 
 		\right\rangle}
 		{\left\langle
-			\left(\sum_{t=0}^h \nabla_\theta \log\pi_\theta(a_t|s_t) \right)^2
+			\left(\sum_{t=0}^T \nabla_\theta \log\pi_\theta(a_t|s_t) \right)^2
 		\right\rangle}
@@
 \begin{equation}
-  \sum_{t=0}^h \nabla_\theta J(\theta) = \mathbb{E}_{\pi_\theta}\left[\sum_{t=0}^h \nabla_\theta{\log\pi_\theta(a|s)}Q_\pi(s,a)\right].
+  \sum_{t=0}^T \nabla_\theta J(\theta) = \mathbb{E}_{\pi_\theta}\left[\sum_{t=0}^T \nabla_\theta{\log\pi_\theta(a|s)}Q_\pi(s,a)\right].
   \label{eqn:egrad}
@@
 \begin{equation}
-  \nabla_\theta J(\theta) \approx \frac{1}{h} \sum_{t=0}^h \nabla_\theta\log\pi_\theta(s) Q(s_t,a_t).
+  \nabla_\theta J(\theta) \approx \frac{1}{T} \sum_{t=0}^T \nabla_\theta\log\pi_\theta(s) Q(s_t,a_t).
 \end{equation}
@@
 
-With these insights in mind, \ref{sec:natural} presents the natural gradient, a refined type of gradient which has a convenient fit in the actor-critic setting we just established.
+With these insights in mind, section \ref{sec:natural} presents the natural gradient, a refined type of gradient which has a convenient fit in the actor-critic setting we just established.
 
@@
 of the natural gradient. 
-The Fisher information matrix represents the certainty we have on our estimate of the gradient and is defined as the covariance of the log likelihood function of a trajectory $\tau_{\pi_\theta}^h$, which as \cite{4863} show can be written as
+The Fisher information matrix represents the certainty we have on our estimate of the gradient and is defined as the covariance of the log likelihood function of a trajectory $\tau_{\pi_\theta}^T$, which as \cite{4863} show can be written as
 \begin{equation}
@@
 \begin{equation}
-  \nabla_\theta J(\theta) = \int_\mathbb{S} {p(s|\theta) \int_\mathbb{A} \pi_\theta(a|s) \nabla_\theta \log \pi(a|s) {\nabla_\theta \log \pi(a|s)}^T dads}\ w \equiv F_\theta w.
+  \nabla_\theta J(\theta) = \int_\mathbb{S} {d^\pi(s) \int_\mathbb{A} \pi_\theta(a|s) \nabla_\theta \log \pi_\theta(a|s) {\nabla_\theta \log \pi_\theta(a|s)}^T dads}\ w \equiv F_\theta w.
   \label{eqn:J-equals-F}
@@
 We show that we can reuse value-estimation approaches in actor-critic settings to improve gradient estimate through better policy evaluation.
-This lead to the introduction of the natural gradient as a way to iterate through policy space instead of parameter space, which improves sample efficiency, especially when the gradient in parameter space is very small, i.e., if there is a plateau.
+This leads to the introduction of the natural gradient as a way to iterate through policy space instead of parameter space, which improves sample efficiency, especially when the gradient in parameter space is very small, i.e., if there is a plateau.
 

### 2019-03-12 pge-mattis.tex
Some improvements

@@
 Given a differentiable model of the policy, we want to apply a gradient-ascent technique to reach the local optimum. 
-Gradient ascent has strong guarantees on convergence, and is theoretically well researched.
+Gradient ascent is theoretically well researched.
 The main issue is that the policy gradient with respect to the expected return is not available, thus we need to estimate it.
-For this reason, the biggest weakness of policy gradient algorithms is sample efficiency, as they tend to require on-policy data for the gradient estimate, and conservative learning rates for the convergence guarantees.
+For this reason, the biggest weakness of policy gradient algorithms is sample efficiency, as they tend to require on-policy data for the gradient estimate.
 Thus, most research is focused on finding algorithms with improved sample efficiency.
-This paper provides a deep introduction to policy gradient that shows the development of policy gradient approaches, and should enable the reader to understand current research in the area.
+This paper provides a deep introduction to policy gradient that shows the development of policy gradient approaches, and should enable the reader to follow current research in the area.
 
@@
 Using a parameterized policy to decide the next action, they can easily incorporate prior domain knowledge, but require a lot of configuration to produce an effective agent for a specific environment.
-Also, they frequently require on-policy training and often converge slower, but can guaranteed find a local optimum given some constraints on the hyperparameters. 
-Options to model a policy gradient method are not only given by the policy model itself, but also in the way a policy improvement is computed. 
+Though they frequently require on-policy training and often converge slower. 
+In this paper, we focus on the computation of policy improvements, though we introduce the most common policy classes shortly.
 In this case, improvement means a step in the parameter space such that the policy under the new parameters will on average perform better than the old policy, i.e. improve its expected return. 
@@
 
-The most frequently used policy classes in policy gradient approaches are Gibbs policies $\pi_\theta(a|s) = \frac{exp(\phi(s,a)^T\theta)}{\sum_b exp(\phi_1(s,b)^T\theta)}$ \cite{Sutton:1999:PGM:3009657.3009806,Bagnell2004LearningD} for discrete problems, and Gaussian policies $\pi_\theta(a|s) = \mathcal{N}(\phi(s,a)^T\theta_1,\theta_2)$ for continuous problems, where $\theta_2$ is an exploration parameter \cite{Williams92simplestatistical,peter:article:1996}, and $\phi(s,a)$ is the vector of basis functions on the state-action pair.
+The most frequently used policy classes in policy gradient approaches are Gibbs policies $\pi_\theta(a|s) = \frac{\exp(\phi(s,a)^T\theta)}{\sum_b \exp(\phi_1(s,b)^T\theta)}$ \cite{Sutton:1999:PGM:3009657.3009806,Bagnell2004LearningD} for discrete problems, and Gaussian policies $\pi_\theta(a|s) = \mathcal{N}(\phi(s,a)^T\theta_1,\theta_2)$ for continuous problems, where $\theta_2$ is an exploration parameter \cite{Williams92simplestatistical,peter:article:1996}, and $\phi(s,a)$ is the vector of basis functions on the state-action pair.
+
+\paragraph{Episodic algorithms.} 
+A very general optimization approach to this optimization problem are episodic algorithms. 
+We take some distribution $p(w)$ over the parameter space of some policy class $\pi_w$, and sample policies from that distribution. 
+The policy class $\pi_w$ is most often chosen deterministic. 
+Using these policies, we perform a large number of roll-outs, and calculate a posterior using the prior $p(w)$ and the returns of our sampled roll-outs. 
+This is repeated until the resulting policies are sufficiently performant, or until $p(w)$ does not change anymore.
+The resulting algorithms are black-box optimizers, and as such are largely applicable, but can not use the problem structure and are often too inefficient.
+Given these insights, we require a way to design algorithms that improve the acting policy directly by observing each interaction with the environment.
 
 \paragraph{Policy gradient.} 
-Our goal with respect to episodes is to maximize the expectation of the total reward, also called expected return. 
+Our goal with respect to episodes is still to maximize the expectation of the total reward, also called expected return. 
 The total reward in the horizon $T$ is $\sum_{t=0}^{T} r_{t}$. 
-We introduce a discount factor $\gamma \in [0,1)$ to trade-off bias ($\gamma\to0$) versus variance ($\gamma\to1$) with a normalization factor of $1-\gamma$, which is often omitted though strongly recommended.
-Intuitively, this reflects the idea that the relevance of later actions declines, which puts more weight on less actions, thus the increased bias for reduced variance. 
+We introduce a discount factor $\gamma \in [0,1)$.
+Intuitively, this reflects the idea that the relevance of later actions declines, and ensures that the return is always finite, even for the infinite horizon $T \to \infty$. 
 The discounted total reward is
@@
 \begin{equation}
-  \mathcal{R}^\tau \equiv \mathcal{R}_0^T := (1-\gamma) \sum_{t=0}^{T} \gamma^t r_t.
+  \mathcal{R}^\tau \equiv \mathcal{R}_0^T := \sum_{t=0}^{T} \gamma^t r_t.
   \label{eqn:acc-reward}
@@
 Thus, we search $\nabla_\theta J(\theta) := \nabla_\theta \mathbb{E}_{p_\pi(\tau)}\left[\mathcal{R}^\tau\right]$, to make a policy gradient step according to $\theta_{k+1} = \theta_k + \alpha_k \nabla_\theta J(\theta)$, where $\alpha_k$ denotes a learning rate.
+\ref{sec:pge} shows how we can estimate $J(\theta)$.
 
@@
 \label{sec:pge}
+
+In this section, we introduce methods for estimating the policy gradient.
 
@@
 However, finite-difference gradients suffer from the curse of dimensionality, and can require very small $\delta\theta$.
-Thus, they only work well in specific scenarios, but should not be discarded due to simplicity.
+When we estimate the gradient this way, we do something similar to the episodic approach, but directly on the acting policy.
+Thus, finite-difference gradients only work well in specific scenarios, but should not be discarded due to simplicity.
 
@@
 For this derivation, we will change the perspective a bit, requiring some additional definitions. 
-We define ${\mu_\pi}_i = \sum_{t=0}^{\infty} p(s_t=s_i | s_0, \pi)$ \cite{puterman2014markov} as the state distribution, though it does not sum up to one without normalization.
-Note that $\mu_\pi$ is equivalent to the discounted state visit count $d^\pi$ introduced by Sutton et al.
-\cite{Sutton:1999:PGM:3009657.3009806}.
+We define ${\mu_\pi}_i = \sum_{t=0}^{\infty} p(s_t=s_i | s_0, \pi)$ \cite{puterman2014markov} as the state distribution, though it does not sum up to one without normalization, which is easily done by multiplying with $(1-\gamma)$.
+Note that $\mu_\pi$ is equivalent to the discounted state visit count $d^\pi$ introduced by Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806}.
 Further, we define $P_\pi$ as the transition matrix, i.e.
 ${P_\pi}_{i,j}=p(s_j|s_i,\pi)$, $r_\pi$ as the mean rewards for all states given by ${r_\pi}_i = \int_\mathbb{A} r(s_i,a)\pi_\theta(a|s_i) da$, and $\mu_0 = \left[\mu_0(s_0),\mu_0(s_1),\ldots\right]^T$ as a vector representing the initial state distribution.
-From which it follows that $V_\pi = \mu_\pi r_\pi$
-From this, we derive
-\begin{align}
-	\mu_\pi &= \mu_0 + P_\pi^T \mu_\pi \nonumber \\
-	(I-P^T)\mu_\pi &= \mu_0 \nonumber \\
-	\mu_\pi^T &= \mu_0^T (I-P_\pi)^{-1} ,
-\end{align}
-so we can reformulate the problem as 
+From which it follows that $V_\pi = \mu_\pi r_\pi$, so we can reformulate the problem as 
 \begin{align}
@@
 \begin{align*}
-	\nabla_\theta J(\theta) &= \nabla_\theta \mu_0^T V_\pi
-	\\ &= \mu_0^T \nabla_\theta \gamma P_\pi V_\pi
-	\\ &= \mu_0^T (I - \gamma P_\pi)^{-1} (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi)
+	\nabla_\theta J(\theta) &= \nabla_\theta \mu_0^T V_\pi = \mu_0^T \nabla_\theta V_\pi
+\end{align*}
+We can replace $\nabla_\theta V_\pi$ using
+\begin{align*}
+	\nabla_\theta V_\pi &= \nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi + \gamma P_\pi \nabla_\theta V_\pi
+	\\ (I-\gamma P_\pi) \nabla_\theta V_\pi &= \nabla_\theta r_\pi + (\nabla_\theta P_\pi) V_\pi
+	\\ \nabla_\theta V_\pi &= (I - \gamma P_\pi)^{-1} (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi), 
+\end{align*}
+and find that
+\begin{align*}
+	\mu_\pi &= \mu_0 + \gamma P_\pi s^T \mu_\pi \nonumber \\
+	(I- \gamma P^T)\mu_\pi &= \mu_0 \nonumber \\
+	\mu_\pi^T &= \mu_0^T (I- \gamma P_\pi)^{-1} ,
+\end{align*}
+which we can take back into the original problem 
+\begin{align*}
+	\nabla_\theta J(\theta) &= \mu_0^T (I - \gamma P_\pi)^{-1} (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi)
 	\\ &= \mu_\pi^T (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) r_\pi)
-	\\ &\equiv \sum\nolimits_{i,j} d^\pi(s_i) \nabla_\theta\pi(a_j|s_i) Q(s_i,a_j)
-	\\ &= \sum\nolimits_{i,j} d^\pi(s_i) \pi(a_j|s_i) \nabla_\theta\log\pi(a_j|s_i) Q_\pi(s_i,a_j).
+	\\ &\equiv \sum\nolimits_{i,j} \mu(s_i) \nabla_\theta\pi(a_j|s_i) Q(s_i,a_j)
+	\\ &= \sum\nolimits_{i,j} \mu(s_i) \pi(a_j|s_i) \nabla_\theta\log\pi(a_j|s_i) Q_\pi(s_i,a_j).
 \end{align*}
@@
 
-\paragraph{Immediate updates.} 
-The first class of algorithms developed to update a policy directly are called REINFORCE \cite{Williams92simplestatistical}.
+The first class of algorithms developed to update a policy directly using a critic are called REINFORCE \cite{Williams92simplestatistical}.
 REINFORCE collects a complete episode, at which point we can calculate the actual state-action value by traversing backwards over the trajectory, and estimates:
@@
 \end{equation}
-where $b_\tau$ is some baseline. 
+This is sometimes also called Monte-Carlo gradient estimation. 
+However, given $r_t > 0, \forall t=0,\dots,h$, we can only increase action probabilities. 
+Obviously, we normalize to ensure $\forall s \in \mathbb{S}: \int_\mathbb{A}{\pi(a|s)da} = 1$. 
+This means actions can only become less probable in relation to other actions. 
+We find that this introduces more variance when learning from samples \cite{Sutton:1999:PGM:3009657.3009806}. 
+One way to counter the variance is to introduce the baseline $b_\tau$.
+We use this baseline to  
 Peters \& Schaal \cite{4867} find that an estimate of the optimal baseline can be calculated by
@@
 
-However, this estimate is very biased.
-To reduce the estimate's bias, we can compute updates from a complete episode.
-
-\paragraph{Episodic updates.} 
-A common way of reducing the bias in value estimation is to sample trajectories, also called roll-outs, and to updates using the complete roll-out. 
-We call these approaches episodic. 
-When we do this, we sum up the immediate updates, which we can rearrange to get
-\begin{equation}
-  \sum_{t=0}^T \nabla_\theta J(\theta) = \mathbb{E}_{\pi_\theta}\left[\sum_{t=0}^T \nabla_\theta{\log\pi_\theta(a|s)}Q_\pi(s,a)\right].
-  \label{eqn:egrad}
-\end{equation}
-
-Episodic REINFORCE \cite{Williams92simplestatistical}, the first approach to this concept, performs roll-outs and takes the final return, at which point we can calculate the actual state-action value by traversing backwards over the trajectory, and estimates
-\begin{equation}
-  \nabla_\theta J(\theta) \approx \frac{1}{T} \sum_{t=0}^T \nabla_\theta\log\pi_\theta(s) Q(s_t,a_t).
-\end{equation}
-This is sometimes also called Monte-Carlo gradient estimation. 
-Given $r_t > 0, \forall t=0,\dots,h$, these approaches can only increase action probabilities. 
-Obviously, we normalize to ensure $\forall s \in \mathbb{S}: \int_\mathbb{A}{\pi(a|s)da} = 1$. 
-This means actions can only become less probable in relation to other actions. 
-We find that this introduces more variance when learning from samples \cite{Sutton:1999:PGM:3009657.3009806}. 
-One approach to counter the variance is to introduce a baseline
-\begin{equation}
-  \hat{Q}(s,a) \approx \gamma^{t} (r_{t} - b(s_t)).
-\end{equation}
-The baseline can represent any prior knowledge or assumption about states. 
-A common way to use this baseline is to use the value function, which gives us the so-called advantage function
+Another common way to use this baseline is to use the value function, which gives us the so-called advantage function
 

### 2019-03-18 pge-mattis.tex
A lot of minor improvements

@@
 
-\title{Policy Gradient Estimation}
-\subtitle{A survey on policy gradient estimators}
+\title{Policy Gradient Algorithms}
+\subtitle{An Introduction}
 
@@
 We define states $s \in \mathbb{S}$, actions $a \in \mathbb{A}$, and rewards $r \in \mathbb{R}$. 
-A trajectory $\tau := (s_0, a_0, $ $s_1, a_1, \dots, s_T, a_T)$ is generated by drawing $s_0 \sim \mu_0(s_0)$ according to the distribution over initial states $\mu_0(s_0)$, and successively sampling $a_t \sim \pi_\theta(a_t|s_t)$ according to the policy $\pi_\theta$, and $s_{t+1} \sim p(s_{t+1}|s_t,a_t)$ until the horizon $T$, or a terminal state is reached.
+A trajectory $\tau := (s_0, a_0, $ $s_1, a_1, \dots, s_T, a_T)$ is generated by drawing $s_0 \sim \mu_0(s_0)$ according to the distribution over initial states $\mu_0(s_0)$, and successively sampling $a_t \sim \pi(a_t|s_t)$ according to the policy $\pi$ parameterized by $\theta$, and $s_{t+1} \sim p(s_{t+1}|s_t,a_t)$ until the horizon $T$, or a terminal state is reached.
 At each time step, we receive a reward according to $r_t = r(s_t, a_t) \equiv \mathbb{E}_{s'}\left[r(s_t,a_t,s')\right]$. 
@@
 
-The most frequently used policy classes in policy gradient approaches are Gibbs policies $\pi_\theta(a|s) = \frac{\exp(\phi(s,a)^T\theta)}{\sum_b \exp(\phi_1(s,b)^T\theta)}$ \cite{Sutton:1999:PGM:3009657.3009806,Bagnell2004LearningD} for discrete problems, and Gaussian policies $\pi_\theta(a|s) = \mathcal{N}(\phi(s,a)^T\theta_1,\theta_2)$ for continuous problems, where $\theta_2$ is an exploration parameter \cite{Williams92simplestatistical,peter:article:1996}, and $\phi(s,a)$ is the vector of basis functions on the state-action pair.
-
-\paragraph{Episodic algorithms.} 
-A very general optimization approach to this optimization problem are episodic algorithms. 
-We take some distribution $p(w)$ over the parameter space of some policy class $\pi_w$, and sample policies from that distribution. 
-The policy class $\pi_w$ is most often chosen deterministic. 
-Using these policies, we perform a large number of roll-outs, and calculate a posterior using the prior $p(w)$ and the returns of our sampled roll-outs. 
-This is repeated until the resulting policies are sufficiently performant, or until $p(w)$ does not change anymore.
-The resulting algorithms are black-box optimizers, and as such are largely applicable, but can not use the problem structure and are often too inefficient.
-Given these insights, we require a way to design algorithms that improve the acting policy directly by observing each interaction with the environment.
+The most frequently used policy classes in policy gradient approaches are Gibbs policies $\pi(a|s) = \frac{\exp(\phi(s,a)^T\theta)}{\sum_b \exp(\phi(s,b)^T\theta)}$ \cite{Sutton:1999:PGM:3009657.3009806,Bagnell2004LearningD} for discrete problems, and Gaussian policies $\pi(a|s) = \mathcal{N}(\phi(s,a)^T\theta_1,\theta_2)$ for continuous problems, where $\theta_2$ is an exploration parameter \cite{Williams92simplestatistical,peter:article:1996}, and $\phi(s,a)$ is the vector of basis functions on the state-action pair.
 
@@
 However, finite-difference gradients suffer from the curse of dimensionality, and can require very small $\delta\theta$.
-When we estimate the gradient this way, we do something similar to the episodic approach, but directly on the acting policy.
 Thus, finite-difference gradients only work well in specific scenarios, but should not be discarded due to simplicity.
@@
 For this derivation, we will change the perspective a bit, requiring some additional definitions. 
-We define ${\mu_\pi}_i = \sum_{t=0}^{\infty} p(s_t=s_i | s_0, \pi)$ \cite{puterman2014markov} as the state distribution, though it does not sum up to one without normalization, which is easily done by multiplying with $(1-\gamma)$.
+We define ${\mu_\pi}_i = \sum_{t=0}^{\infty} \gamma^t p(s_t=s_i | s_0, \pi)$ \cite{puterman2014markov} as the state distribution, though it does not sum up to one without normalization, which is easily done by multiplying with $(1-\gamma)$.
 Note that $\mu_\pi$ is equivalent to the discounted state visit count $d^\pi$ introduced by Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806}.
 Further, we define $P_\pi$ as the transition matrix, i.e.
-${P_\pi}_{i,j}=p(s_j|s_i,\pi)$, $r_\pi$ as the mean rewards for all states given by ${r_\pi}_i = \int_\mathbb{A} r(s_i,a)\pi_\theta(a|s_i) da$, and $\mu_0 = \left[\mu_0(s_0),\mu_0(s_1),\ldots\right]^T$ as a vector representing the initial state distribution.
+${P_\pi}_{i,j}=\sum\nolimits_k p(s_j|s_i,a_k)\pi(a_k|s_i)$, $r_\pi$ as the mean rewards for all states given by ${r_\pi}_i = \sum_j r(s_i,a_j)\pi(a_j|s_i)$, and $\mu_0 = \left[\mu_0(s_0),\mu_0(s_1),\ldots\right]^T$ as a vector representing the initial state distribution.
 From which it follows that $V_\pi = \mu_\pi r_\pi$, so we can reformulate the problem as 
@@
 \end{align}
-Where $V_\pi = \left[ V_\pi(s_0), V_\pi(s_1), \ldots \right]^T$. We derive 
+Where $V_\pi = \left[ V_\pi(s_0), V_\pi(s_1), \ldots \right]^T$. 
+We derive 
 \begin{align*}
@@
 \begin{align*}
-	\mu_\pi &= \mu_0 + \gamma P_\pi s^T \mu_\pi \nonumber \\
-	(I- \gamma P^T)\mu_\pi &= \mu_0 \nonumber \\
+	\mu_\pi &= \mu_0 + \gamma P_\pi^T \mu_\pi \nonumber \\
+	(I- \gamma P_\pi^T)\mu_\pi &= \mu_0 \nonumber \\
 	\mu_\pi^T &= \mu_0^T (I- \gamma P_\pi)^{-1} ,
 \end{align*}
-which we can take back into the original problem 
-\begin{align*}
-	\nabla_\theta J(\theta) &= \mu_0^T (I - \gamma P_\pi)^{-1} (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi)
-	\\ &= \mu_\pi^T (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) r_\pi)
-	\\ &\equiv \sum\nolimits_{i,j} \mu(s_i) \nabla_\theta\pi(a_j|s_i) Q(s_i,a_j)
-	\\ &= \sum\nolimits_{i,j} \mu(s_i) \pi(a_j|s_i) \nabla_\theta\log\pi(a_j|s_i) Q_\pi(s_i,a_j).
-\end{align*}
-Where we used $\nabla_\theta \pi_\theta(a|s) = \pi_\theta(a|s)\nabla_\theta\log\pi_\theta(a|s)$, obtained from the likelihood ratio, which gives us the likelihood-ratio gradient,
+which we can take back into the gradient equation 
+\begin{align}
+	\nabla_\theta J(\theta) &= \mu_0^T (I - \gamma P_\pi)^{-1} (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi) \nonumber
+	\\ &= \mu_\pi^T (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi) \nonumber
+	\\ &\equiv \sum\nolimits_{i,j} \mu(s_i) \nabla_\theta\pi(a_j|s_i) Q_\pi(s_i,a_j) \label{eqn:likelihood-equiv}
+	\\ &= \sum\nolimits_{i,j} \mu(s_i) \pi(a_j|s_i) \nabla_\theta\log\pi(a_j|s_i) Q_\pi(s_i,a_j). \nonumber
+\end{align}
+
+The equivalence in \eqref{eqn:likelihood-equiv} comes from the observation that $\nabla_\theta \pi(a|s)$ is distributed over the addition $\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi$.
+When we take out this common factor, $Q_\pi(s,a)$ remains.
+Then, we used $\nabla_\theta \pi(a|s) = \pi(a|s)\nabla_\theta\log\pi(a|s)$, obtained from the likelihood ratio $\nabla_\theta \log p(x|\theta) = \frac{\nabla_\theta p(x|\theta)}{p(x|\theta)}$. 
+This gives us the likelihood-ratio gradient,
 \begin{equation}
-	\nabla_\theta J(\theta) = \mathbb{E}_{\substack{\ s \sim d^\pi \\a \sim \pi}} \Big[\nabla_\theta{\log\pi_\theta(a|s)}Q_\pi(s,a)\Big].
+	\nabla_\theta J(\theta) = \mathbb{E}_{\substack{\ s \sim \mu_\pi \\a \sim \pi}} \Big[\nabla_\theta{\log\pi(a|s)}Q_\pi(s,a)\Big].
 	\label{eqn:like-grad}
@@
 
+\paragraph{Episode-based updates.} 
+A very general optimization approach to this optimization problem are episodic algorithms. 
+We take a search distribution $p(\theta|\omega)$ over the parameter space of the policy class $\pi$ of the acting policy, and sample policies from that distribution. 
+The policy class $\pi$ is most often chosen deterministic. 
+Using these policies, we sample trajectories $\tau$, and update the search policy using the returns of our sampled roll-outs
+\begin{equation}
+	\nabla_\theta J(\theta) \approx \sum_{t=0}^T \nabla_\omega \log p(\theta|\omega) \mathcal{R}_t^T.
+\end{equation}
+The resulting algorithms are black-box optimizers, and as such are largely applicable, but can not use the problem structure and have a lot of variance.
+Given these insights, we require a way to design algorithms that improve the acting policy stepwise by observing each interaction with the environment.
+
+\paragraph{Step-based updates.}
 The first class of algorithms developed to update a policy directly using a critic are called REINFORCE \cite{Williams92simplestatistical}.
-REINFORCE collects a complete episode, at which point we can calculate the actual state-action value by traversing backwards over the trajectory, and estimates:
+REINFORCE samples a complete episode, at which point we can calculate the actual state-action value by traversing backwards over the trajectory, and estimates:
 \begin{equation}
-	\nabla_\theta J(\theta) \approx \nabla_\theta\log\pi_\theta(s) \left(Q(s_t,a_t) - b_\tau \right).
+	\nabla_\theta J(\theta) \approx \nabla_\theta\log\pi(s) \left(Q(s_t,a_t) - b_\tau \right).
 	\label{eqn:reinforce}
@@
 One way to counter the variance is to introduce the baseline $b_\tau$.
-We use this baseline to  
-Peters \& Schaal \cite{4867} find that an estimate of the optimal baseline can be calculated by
+Peters et al. \cite{4867} find that an estimate of the optimal baseline can be calculated by
 \begin{equation}
@@
 		{\left\langle 
-			\left(\sum_{t=0}^T \nabla_\theta \log\pi_\theta(a_t|s_t) \right)^2 \sum_{t'=0}^T a_{t'} r_{t'} 
+			\left(\sum_{t=0}^T \nabla_\theta \log\pi(a_t|s_t) \right)^2 \sum_{t'=0}^T a_{t'} r_{t'} 
 		\right\rangle}
 		{\left\langle
-			\left(\sum_{t=0}^T \nabla_\theta \log\pi_\theta(a_t|s_t) \right)^2
+			\left(\sum_{t=0}^T \nabla_\theta \log\pi(a_t|s_t) \right)^2
 		\right\rangle}
@@
 
+Policy gradient methods can be described in terms of two main steps often called policy evaluation and policy improvement. 
+For actor-critic approaches, we separate these steps from the actor component by implementing a critic. 
+This means, the actor consists only of the policy, while the critic is focused on estimating a score for the actions taken. 
+By that concept, observations of the environment are given to the actor only to decide the next action, and to the critic only to improve its function estimation with the respective rewards. 
+Figure \ref{fig:ac} shows the general structure of an actor-critic algorithm.
+
 \tikzset{block/.style= {draw, rectangle, align=center, minimum height=2em, minimum width=3cm}}
@@
     \node [block, name=env, minimum width=6cm] (env) {Environment};
-    \node [block, above of=env, node distance=1.5cm] (critic) {Critic \\ $\hat{Q}^\pi(s,a)$};
+    \node [block, above of=env, node distance=1.5cm] (critic) {Critic \\ $\hat{Q}_\pi(s,a)$};
     \node [block, above of=critic] (actor) {Actor \\ policy $\pi(a|s)$};
@@
   \end{tikzpicture}
-  \caption{An actor-critic framework comparable to Kimura et al.  \cite{Kimura1998AnAO}.} \label{fig:ac}
+  \caption{A visualization inspired by Kimura et al. \cite{Kimura1998AnAO}, showing the actor-critic framework.} \label{fig:ac}
 \end{figure}
-
-Policy gradient methods can be described in terms of two main steps often called policy evaluation and policy improvement. 
-For actor-critic approaches, we separate these steps from the actor component by implementing a critic. 
-This means, the actor consists only of the policy, while the critic is focused on estimating a score for the actions taken. 
-By that concept, observations of the environment are given to the actor only to decide the next action, and to the critic only to improve its function estimation with the respective rewards. 
-Figure \ref{fig:ac} shows the general structure of an actor-critic algorithm.
 
@@
 \begin{equation}
-  r_t + \gamma \hat{V}_\pi(s_{t+1}) - \hat{V}_\pi(s_t).
+  \delta(s_t) = r_t + \gamma \hat{V}_\pi(s_{t+1}) - \hat{V}_\pi(s_t).
 \end{equation}
@@
   \includegraphics[width=\textwidth]{nat-grad-adv}
-  \caption{An experiment showing where the natural gradient has a great advantage. \cite{Peters_IICHR_2003} }\label{fig:nat-grad-adv}
+  \caption{An experiment showing where the natural gradient has a great advantage \cite{Peters_IICHR_2003}. }\label{fig:nat-grad-adv}
 \end{figure}
@@
 of the natural gradient. 
-The Fisher information matrix represents the certainty we have on our estimate of the gradient and is defined as the covariance of the log likelihood function of a trajectory $\tau_{\pi_\theta}^T$, which as \cite{4863} show can be written as
+The Fisher information matrix represents the certainty we have on our estimate of the gradient and is defined as the covariance of the log likelihood function of a trajectory $\tau_{\pi}^T$, which as \cite{4863} show can be written as
 \begin{equation}
-  F_\theta = \int_\mathbb{S} d^\pi(s) \int_\mathbb{A} \pi_\theta(a|s) \nabla_\theta \log{\pi_\theta(a|s)} \nabla_\theta \log{\pi_\theta(a|s)}^T dads.
+  F_\theta = \int_\mathbb{S} d^\pi(s) \int_\mathbb{A} \pi(a|s) \nabla_\theta \log{\pi(a|s)} \nabla_\theta \log{\pi(a|s)}^T dads.
   \label{eqn:F}
@@
 \begin{equation}
-  \nabla_\theta J(\theta) = \int_\mathbb{S} {d^\pi(s) \int_\mathbb{A} \pi_\theta(a|s) \nabla_\theta \log \pi_\theta(a|s) {\nabla_\theta \log \pi_\theta(a|s)}^T dads}\ w \equiv F_\theta w.
+  \nabla_\theta J(\theta) = \int_\mathbb{S} {d^\pi(s) \int_\mathbb{A} \pi(a|s) \nabla_\theta \log \pi(a|s) {\nabla_\theta \log \pi(a|s)}^T dads}\ w \equiv F_\theta w.
   \label{eqn:J-equals-F}
 \end{equation}
+Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806} call this estimator compatible.
+
 From \eqref{eqn:nat-grad}, \eqref{eqn:F}, and \eqref{eqn:J-equals-F}, it follows that

### 2019-03-18 pge-mattis.tex
Clean up a sentence

@@
 A very general optimization approach to this optimization problem are episodic algorithms. 
-We take a search distribution $p(\theta|\omega)$ over the parameter space of the policy class $\pi$ of the acting policy, and sample policies from that distribution. 
+We take a search distribution $p(\theta|\omega)$ over the parameter space of the policy class $\pi$, and sample acting policies from that distribution. 
 The policy class $\pi$ is most often chosen deterministic. 

### 2019-03-20 pge-mattis.tex
Last fixes before submission

@@
 \RequirePackage{fix-cm}
-\documentclass[draft]{svjour3}
+\documentclass[final]{svjour3}
 \smartqed
@@
 \usepackage[utf8]{inputenc}
-\usepackage[final]{graphicx}
+\usepackage{graphicx}
 \graphicspath{ {./images/} }
 \usepackage{amsmath,amssymb}
-\usepackage[final]{hyperref}
+\usepackage{hyperref}
 \usepackage{tikz}
@@
 Where we define $i:j$ with $i,j \in \mathbb{N}, i < j$ as an index over all integers from $i$ to $j$, i.e., $s_{i:j} \equiv s_i, s_{i+1}, \dots, s_j$. 
-We assume no additional prior knowledge about the environment, thus we assume the probability of a trajectory is 
+We assume no additional prior knowledge about the environment, meaning we assume the probability of a trajectory is 
 \begin{equation}
@@
 \paragraph{Policy gradient.} 
-Our goal with respect to episodes is still to maximize the expectation of the total reward, also called expected return. 
+Our goal with respect to episodes is to maximize the expectation of the total reward, also called expected return. 
 The total reward in the horizon $T$ is $\sum_{t=0}^{T} r_{t}$. 
 We introduce a discount factor $\gamma \in [0,1)$.
-Intuitively, this reflects the idea that the relevance of later actions declines, and ensures that the return is always finite, even for the infinite horizon $T \to \infty$. 
+Intuitively, this reflects the idea that the relevance of later actions declines, and ensures that the return is finite, even for the infinite horizon $T \to \infty$. 
 The discounted total reward is 
-
 \begin{equation}
@@
 Thus, we search $\nabla_\theta J(\theta) := \nabla_\theta \mathbb{E}_{p_\pi(\tau)}\left[\mathcal{R}^\tau\right]$, to make a policy gradient step according to $\theta_{k+1} = \theta_k + \alpha_k \nabla_\theta J(\theta)$, where $\alpha_k$ denotes a learning rate.
-\ref{sec:pge} shows how we can estimate $J(\theta)$.
+Section \ref{sec:pge} shows how we can estimate $J(\theta)$.
 
@@
 \end{equation}
-
-As we will see, this function also gives us the true gradient of $J(\theta)$, though in general we need to estimate it. 
 In addition to the value function, we also define the state-action value function, often called Q-function. 
@@
 \end{equation}
-
+As we will see, this function also gives us the true gradient of $J(\theta)$, though in general we need to estimate it. 
 Using the value function, and the Q-function, we can derive a better estimate of the policy gradient.
@@
 For this derivation, we will change the perspective a bit, requiring some additional definitions. 
-We define ${\mu_\pi}_i = \sum_{t=0}^{\infty} \gamma^t p(s_t=s_i | s_0, \pi)$ \cite{puterman2014markov} as the state distribution, though it does not sum up to one without normalization, which is easily done by multiplying with $(1-\gamma)$.
+We define ${\mu_\pi}_i = \sum_{t=0}^{\infty} \gamma^t p(s_t=s_i | s_0, \pi)$ \cite{puterman2014markov} as the discounted state distribution, though it does not sum up to one without normalization, which can be achieved by multiplying with $(1-\gamma)$.
 Note that $\mu_\pi$ is equivalent to the discounted state visit count $d^\pi$ introduced by Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806}.
@@
 ${P_\pi}_{i,j}=\sum\nolimits_k p(s_j|s_i,a_k)\pi(a_k|s_i)$, $r_\pi$ as the mean rewards for all states given by ${r_\pi}_i = \sum_j r(s_i,a_j)\pi(a_j|s_i)$, and $\mu_0 = \left[\mu_0(s_0),\mu_0(s_1),\ldots\right]^T$ as a vector representing the initial state distribution.
-From which it follows that $V_\pi = \mu_\pi r_\pi$, so we can reformulate the problem as 
+Finally, we define $V_\pi = \left[ V_\pi(s_0), V_\pi(s_1), \ldots \right]^T$, from which it follows that $V_\pi = \mu_\pi r_\pi$, so we can reformulate the problem as 
 \begin{align}
@@
 \end{align}
-Where $V_\pi = \left[ V_\pi(s_0), V_\pi(s_1), \ldots \right]^T$. 
 We derive 
@@
 Then, we used $\nabla_\theta \pi(a|s) = \pi(a|s)\nabla_\theta\log\pi(a|s)$, obtained from the likelihood ratio $\nabla_\theta \log p(x|\theta) = \frac{\nabla_\theta p(x|\theta)}{p(x|\theta)}$. 
-This gives us the likelihood-ratio gradient,
+This gives us the likelihood-ratio gradient
 \begin{equation}
-	\nabla_\theta J(\theta) = \mathbb{E}_{\substack{\ s \sim \mu_\pi \\a \sim \pi}} \Big[\nabla_\theta{\log\pi(a|s)}Q_\pi(s,a)\Big].
+	\nabla_\theta J(\theta) = \mathbb{E}_{\substack{\ s \sim \mu_\pi \\a \sim \pi}} \Big[\nabla_\theta{\log\pi(a|s)}Q_\pi(s,a)\Big],
 	\label{eqn:like-grad}
@@
 intuitively meaning we should to increase the probability of actions that lead to higher Q-values.
+This formulation enables us to calculate the gradient $\nabla_\theta J(\theta)$, while directly taking advantage of the MDP structure in the form of $Q_\pi(s,a)$.
 
 Obviously, we do not have the true $Q_\pi(s,a)$, thus we need to approximate it by $\hat{Q}_\pi(s,a)$. 
-We draw $a \sim \pi(a|s)$, starting in state $s_0 \sim p(s_0)$ and match a function estimator to our observations.
-It is shown that for $\lim_{k\to\infty}\alpha_k = 0$, and $\sum_{k=0}^\infty \alpha_k$ we are guaranteed to converge to a local optimum \cite{Sutton:1999:PGM:3009657.3009806}.
+In all of the following sections, when we say that we sample an episode, we mean to draw $a \sim \pi(a|s)$, starting in state $s_0 \sim p(s_0)$ and match a function estimator to our observations.
+Following this procedure, it is shown that for $\lim_{k\to\infty}\alpha_k = 0$, and $\sum_{k=0}^\infty \alpha_k$ we are guaranteed to converge to a local optimum \cite{Sutton:1999:PGM:3009657.3009806}.
 Approximating $Q_\pi(s,a)$ by an unbiased estimator $f_w(s_t, a_t) \equiv \hat{Q}_\pi(s_t, a_t)$, Sutton et al.
-\cite{Sutton:1999:PGM:3009657.3009806} show that using this function approximation we will converge to the real local optimum.
+\cite{Sutton:1999:PGM:3009657.3009806} show that using this function approximation we will converge to the true local optimum of $J(\theta)$.
 
@@
 \end{equation}
-The resulting algorithms are black-box optimizers, and as such are largely applicable, but can not use the problem structure and have a lot of variance.
+The resulting algorithms are black-box optimizers, and as such are largely applicable, but can not use any temporal information and have a lot of variance.
 Given these insights, we require a way to design algorithms that improve the acting policy stepwise by observing each interaction with the environment.
@@
 This means actions can only become less probable in relation to other actions. 
-We find that this introduces more variance when learning from samples \cite{Sutton:1999:PGM:3009657.3009806}. 
+We find that this introduces more variance when learning from samples \cite{Sutton:1999:PGM:3009657.3009806}, and by that defeats the purpose of why we thought of this approach in the first place. 
 One way to counter the variance is to introduce the baseline $b_\tau$.
@@
 			\left(\sum_{t=0}^T \nabla_\theta \log\pi(a_t|s_t) \right)^2
-		\right\rangle}
+		\right\rangle},
 \end{equation}
+which does not affect the unbiasedness of the estimate.
 
@@
 Figure \ref{fig:ac} shows the general structure of an actor-critic algorithm.
+Given this definition, we can already say that the algorithms presented at the end of section \ref{sec:pge} are actor-critic approaches.
 
@@
 \begin{figure}
+  \begin{center}
   \begin{tikzpicture}[auto, node distance=2cm,>=latex']
@@
   \end{tikzpicture}
+  \end{center}
   \caption{A visualization inspired by Kimura et al. \cite{Kimura1998AnAO}, showing the actor-critic framework.} \label{fig:ac}
@@
 The critic estimates a state-action value function as defined in \eqref{eqn:q}. 
-\cite{Sutton:1999:PGM:3009657.3009806}, and \cite{Konda:2003:AA:942271.942292} find that the estimation $f_w^\pi(s,a) \approx Q_\pi(s,a)$ does not affect the unbiasedness of the gradient estimate under some restrictions.
+Sutton et al.
+\cite{Sutton:1999:PGM:3009657.3009806}, and Konda et al.
+\cite{Konda:2003:AA:942271.942292} find that the estimation $f_w^\pi(s,a) \approx Q_\pi(s,a)$ does not affect the unbiasedness of the gradient estimate under some restrictions.
 Specifically, this holds for $f_w^\pi(s,a) = {\nabla_\theta \log\pi(a|s)}^T w$, thus $f_w^\pi(s,a)$ being a linear function parameterized by the vector $w$.
+Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806} call this estimator compatible.
 This guarantees that the function estimator does not cause divergence, and really enables recent research in reinforcement learning for continuous control problems, e.g., in humanoid robotics.
@@
   \delta(s_t) = r_t + \gamma \hat{V}_\pi(s_{t+1}) - \hat{V}_\pi(s_t).
+  \label{eqn:td-error}
 \end{equation}
 
-However, Sutton et al. \cite{1993b} find that this is only guaranteed to be unbiased, if $\int_\mathbb{S}{\pi(s,a)f_w(s,a)da} = 0$ for each state. 
+However, Sutton et al. \cite{1993b} find that this is only guaranteed to be unbiased, if $\int_\mathbb{A}{\pi(s,a)f_w(s,a)da} = 0$ for each state. 
 Given this assumption, the function estimator $f_w$ is limited to an advantage function \eqref{eqn:adv}, which requires bootstrapping for $V_\pi(s,a)$. 
-If we use temporal difference in this context, we run into a problem, as \eqref{eqn:adv} subtracts $\hat{V}_\pi(s_t)$, meaning we would only learn immediate rewards \cite{Peters_IICHR_2003}. 
+If we use temporal difference in this context, we run into a problem, as \eqref{eqn:td-error} subtracts $\hat{V}_\pi(s_t)$, meaning we would only learn immediate rewards \cite{Peters_IICHR_2003}. 
 This would render the process biased. 
@@
 \begin{align}
-  Q_\pi(s,a) = A_\pi(s,a) + V_\pi(s) &= r(s,a) + \gamma \int_\mathbb{S} p(s'|s,a)V_\pi(s')ds' \\
-  \nabla_\theta \log \pi(a_t|s_t)^T w + \phi(s_t)^T v &= r(s_t,a_t) + \gamma \phi(s_{t+1})^T v + \epsilon(s_t,a_t,s_{t+1})
+  Q_\pi(s,a) = A_\pi(s,a) + V_\pi(s) &= r(s,a) + \gamma \int_\mathbb{S} p(s'|s,a)V_\pi(s')ds', \\
+  \nabla_\theta \log \pi(a_t|s_t)^T w + \phi(s_t)^T v &= r(s_t,a_t) + \gamma \phi(s_{t+1})^T v + \epsilon(s_t,a_t,s_{t+1}).
 \end{align}
@@
 \end{equation}
-Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806} call this estimator compatible.
 
@@
 We show that we can reuse value-estimation approaches in actor-critic settings to improve gradient estimate through better policy evaluation.
-This leads to the introduction of the natural gradient as a way to iterate through policy space instead of parameter space, which improves sample efficiency, especially when the gradient in parameter space is very small, i.e., if there is a plateau.
+This leads to the introduction of the natural gradient as a way to iterate through policy space instead of parameter space, which improves sample efficiency, especially when the gradient in parameter space is very small.
 

### 2019-03-20 pge-mattis.tex
Clarify derivation
Yet more minor improvements

@@
 Using our insights from section \ref{sec:pge}, we derive the actor-critic framework in section \ref{sec:ac}, which harnesses value-function estimation for improved gradient updates. 
-Then, in section \ref{sec:natural}, we introduce some gradient ascent methods that build on the approaches given in \ref{sec:pge}, refining the gradient estimation by Fisher's information matrix to get the natural gradient.
+Then, in section \ref{sec:natural}, we introduce some gradient-ascent methods that build on the approaches given in \ref{sec:pge}, refining the gradient estimation by Fisher's information matrix to get the natural gradient.
 Finally, in section \ref{sec:outro}, we give a conclusion of the reviewed approaches, and discuss potential areas of improvement.
@@
 \end{align}
-We derive 
+Since $\mu_0^T$ does not depend on $\theta$, it is obvious that 
 \begin{align*}
-	\nabla_\theta J(\theta) &= \nabla_\theta \mu_0^T V_\pi = \mu_0^T \nabla_\theta V_\pi
+	\nabla_\theta J(\theta) &= \nabla_\theta \mu_0^T V_\pi = \mu_0^T \nabla_\theta V_\pi.
 \end{align*}
@@
 \begin{align*}
-	\nabla_\theta V_\pi &= \nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi + \gamma P_\pi \nabla_\theta V_\pi
-	\\ (I-\gamma P_\pi) \nabla_\theta V_\pi &= \nabla_\theta r_\pi + (\nabla_\theta P_\pi) V_\pi
+	\nabla_\theta V_\pi &= \nabla_\theta \left( r_\pi + \gamma P_\pi V_\pi \right)
+	\\ \nabla_\theta V_\pi &= \nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi + \gamma P_\pi \nabla_\theta V_\pi
+	\\ (I-\gamma P_\pi) \nabla_\theta V_\pi &= \nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi
 	\\ \nabla_\theta V_\pi &= (I - \gamma P_\pi)^{-1} (\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi), 
@@
 \end{align}
-
 The equivalence in \eqref{eqn:likelihood-equiv} comes from the observation that $\nabla_\theta \pi(a|s)$ is distributed over the addition $\nabla_\theta r_\pi + \gamma (\nabla_\theta P_\pi) V_\pi$.
@@
 \cite{Konda:2003:AA:942271.942292} find that the estimation $f_w^\pi(s,a) \approx Q_\pi(s,a)$ does not affect the unbiasedness of the gradient estimate under some restrictions.
-Specifically, this holds for $f_w^\pi(s,a) = {\nabla_\theta \log\pi(a|s)}^T w$, thus $f_w^\pi(s,a)$ being a linear function parameterized by the vector $w$.
+Specifically, this holds for 
+\begin{equation}
+	f_w^\pi(s,a) = {\nabla_\theta \log\pi(a|s)}^T w,
+	\label{eqn:compatible-approximator}
+\end{equation}
+thus $f_w^\pi(s,a)$ being a linear function parameterized by the vector $w$.
 Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806} call this estimator compatible.
@@
 
-However, Sutton et al. \cite{1993b} find that this is only guaranteed to be unbiased, if $\int_\mathbb{A}{\pi(s,a)f_w(s,a)da} = 0$ for each state. 
+However, Sutton et al. \cite{1993b} find that this is only guaranteed to be unbiased, if $\int_\mathbb{A}{\pi(s,a)f_w(s,a)da} = 0, \forall s \in \mathbb{S}$. 
 Given this assumption, the function estimator $f_w$ is limited to an advantage function \eqref{eqn:adv}, which requires bootstrapping for $V_\pi(s,a)$. 
@@
 \end{align}
-which gives involves only linear equations to solve. \cite{4863}
+which gives involves only linear equations to solve \cite{4863}.
 
@@
 Using a value function estimator and calculating the natural gradient, we get the natural policy gradient algorithm (NPG) \cite{NIPS2017_7233}.
-But, if we recall the definition \eqref{eqn:like-grad} of likelihood-ratio gradients, we see that setting $\hat{Q}(s,a) := \nabla_\theta \log \pi(a|s)^T w$, we get
+But, if we recall the definition \eqref{eqn:like-grad} of likelihood-ratio gradients, and the compatible function approximator from \eqref{eqn:compatible-approximator}, we get
 \begin{equation}

### 2019-03-20 pge-mattis.tex
Improve formulation of the TD bootstrapping problem

@@
 The first class of algorithms developed to update a policy directly using a critic are called REINFORCE \cite{Williams92simplestatistical}.
-REINFORCE samples a complete episode, at which point we can calculate the actual state-action value by traversing backwards over the trajectory, and estimates:
+REINFORCE samples a complete episode, at which point we can calculate the actual state-action value by traversing backwards over the trajectory, and estimates
 \begin{equation}
@@
 This is sometimes also called Monte-Carlo gradient estimation. 
-However, given $r_t > 0, \forall t=0,\dots,h$, we can only increase action probabilities. 
+However, given $b_\tau = 0, r_t > 0, \forall t=0,\dots,h$, we can only increase action probabilities. 
 Obviously, we normalize to ensure $\forall s \in \mathbb{S}: \int_\mathbb{A}{\pi(a|s)da} = 1$. 
@@
 We find that this introduces more variance when learning from samples \cite{Sutton:1999:PGM:3009657.3009806}, and by that defeats the purpose of why we thought of this approach in the first place. 
-One way to counter the variance is to introduce the baseline $b_\tau$.
+One way to counter the variance is to use an effective baseline $b_\tau$.
 Peters et al. \cite{4867} find that an estimate of the optimal baseline can be calculated by
@@
 which does not affect the unbiasedness of the estimate.
-
-Another common way to use this baseline is to use the value function, which gives us the so-called advantage function
-
-\begin{equation}
-	A_{\pi}(s_t, a_t) := Q_{\pi}(s_t, a_t) - V_{\pi}(s_t).
-	\label{eqn:adv}
-\end{equation}
 
@@
 However, Sutton et al. \cite{1993b} find that this is only guaranteed to be unbiased, if $\int_\mathbb{A}{\pi(s,a)f_w(s,a)da} = 0, \forall s \in \mathbb{S}$. 
-Given this assumption, the function estimator $f_w$ is limited to an advantage function \eqref{eqn:adv}, which requires bootstrapping for $V_\pi(s,a)$. 
-If we use temporal difference in this context, we run into a problem, as \eqref{eqn:td-error} subtracts $\hat{V}_\pi(s_t)$, meaning we would only learn immediate rewards \cite{Peters_IICHR_2003}. 
+Given this assumption, the function estimator $f_w$ is limited to approximating an advantage function
+\begin{equation}
+	f_w(s_t, a_t) \equiv \hat{A}_{\pi}(s_t, a_t) = \hat{Q}_{\pi}(s_t, a_t) - \hat{V}_{\pi}(s_t),
+	\label{eqn:adv}
+\end{equation} 
+which requires bootstrapping for $\hat{V}_\pi(s)$. 
+If we use temporal difference in this context, we run into a problem, as \eqref{eqn:adv} subtracts $\hat{V}_\pi(s_t)$, meaning we would only learn immediate rewards \cite{Peters_IICHR_2003}. 
 This would render the process biased. 

### 2019-03-20 pge-mattis.tex
Improvements in actor-critic section

@@
 Following this procedure, it is shown that for $\lim_{k\to\infty}\alpha_k = 0$, and $\sum_{k=0}^\infty \alpha_k$ we are guaranteed to converge to a local optimum \cite{Sutton:1999:PGM:3009657.3009806}.
-Approximating $Q_\pi(s,a)$ by an unbiased estimator $f_w(s_t, a_t) \equiv \hat{Q}_\pi(s_t, a_t)$, Sutton et al.
+Approximating $Q_\pi(s,a)$ by an unbiased estimator $f_w^\pi(s_t, a_t) \equiv \hat{Q}_\pi(s_t, a_t)$, Sutton et al.
 \cite{Sutton:1999:PGM:3009657.3009806} show that using this function approximation we will converge to the true local optimum of $J(\theta)$.
@@
 
-However, Sutton et al. \cite{1993b} find that this is only guaranteed to be unbiased, if $\int_\mathbb{A}{\pi(s,a)f_w(s,a)da} = 0, \forall s \in \mathbb{S}$. 
-Given this assumption, the function estimator $f_w$ is limited to approximating an advantage function
+However, Sutton et al. \cite{1993b} find that this is only guaranteed to be unbiased, if $\int_\mathbb{A}{\pi(s,a)f_w^\pi(s,a)da} = 0, \forall s \in \mathbb{S}$. 
+Given this assumption, the function estimator $f_w^\pi$ is limited to approximating an advantage function
 \begin{equation}
-	f_w(s_t, a_t) \equiv \hat{A}_{\pi}(s_t, a_t) = \hat{Q}_{\pi}(s_t, a_t) - \hat{V}_{\pi}(s_t),
+	f_w^\pi(s_t, a_t) \equiv \hat{A}_{\pi}(s_t, a_t) = \hat{Q}_{\pi}(s_t, a_t) - \hat{V}_{\pi}(s_t),
 	\label{eqn:adv}
 \end{equation} 
-which requires bootstrapping for $\hat{V}_\pi(s)$. 
+which requires bootstrapping for $\hat{V}_\pi$. 
 If we use temporal difference in this context, we run into a problem, as \eqref{eqn:adv} subtracts $\hat{V}_\pi(s_t)$, meaning we would only learn immediate rewards \cite{Peters_IICHR_2003}. 
@@
 However, Peters et al. \cite{4863} find that this approximation is highly reliant on the distribution of the training data. 
-This comes from the realization, that we use only a subspace of the actual action value function in $f_w^\pi$, which is only a state value function. 
+This comes from the realization, that we use only a subspace of the actual action value function in $\hat{V}^\pi$, which is only a state value function. 
 One can compare this to approximating a parabola by a line, whereby the approximation changes wildly depending on which part of the parabola is in the training data. 
-An approach to solve this bootstrapping problem is to rewrite the Bellman Equation using \eqref{eqn:adv} and \eqref{eqn:q} with $A_\pi(s,a) = f_w^\pi(s,a)$, $V_\pi(s) = \phi(s)^T v$, and a zero-mean error term $\epsilon(s_t,a_t,s_{t+1})$, we get
+An approach to solve this bootstrapping problem is to rewrite the Bellman Equation using \eqref{eqn:adv} and \eqref{eqn:q} with $\hat{A}_\pi(s,a) = f_w^\pi(s,a)$, $\hat{V}_\pi(s) = \phi(s)^T v$, and a zero-mean error term $\epsilon(s_t,a_t,s_{t+1})$, we get
 \begin{align}
-  Q_\pi(s,a) = A_\pi(s,a) + V_\pi(s) &= r(s,a) + \gamma \int_\mathbb{S} p(s'|s,a)V_\pi(s')ds', \\
+  \hat{Q}_\pi(s,a) = \hat{A}_\pi(s,a) + \hat{V}_\pi(s) &= r(s,a) + \gamma \int_\mathbb{S} p(s'|s,a)\hat{V}_\pi(s')ds', \\
   \nabla_\theta \log \pi(a_t|s_t)^T w + \phi(s_t)^T v &= r(s_t,a_t) + \gamma \phi(s_{t+1})^T v + \epsilon(s_t,a_t,s_{t+1}).
@@
 
-From \eqref{eqn:nat-grad}, \eqref{eqn:F}, and \eqref{eqn:J-equals-F}, it follows that
+From \eqref{eqn:nat-grad}, and \eqref{eqn:J-equals-F}, it follows that
 \begin{equation}
@@
 Peters et al.
-\cite{4863} present this idea and suggest LSTD($\lambda$)-Q, a version of least-squares temporal difference learning  \cite{Boyan:1999:LTD:645528.657618}, as well as episodic natural actor-critic (eNAC).
+\cite{4863} present this idea and suggest LSTD-Q($\lambda$), a version of least-squares temporal difference learning  \cite{Boyan:1999:LTD:645528.657618}, as well as episodic natural actor-critic (eNAC).
 

### 2019-03-20 pge-mattis.tex
Remove misleading citation

@@
 For this derivation, we will change the perspective a bit, requiring some additional definitions. 
-We define ${\mu_\pi}_i = \sum_{t=0}^{\infty} \gamma^t p(s_t=s_i | s_0, \pi)$ \cite{puterman2014markov} as the discounted state distribution, though it does not sum up to one without normalization, which can be achieved by multiplying with $(1-\gamma)$.
+We define ${\mu_\pi}_i = \sum_{t=0}^{\infty} \gamma^t p(s_t=s_i | s_0, \pi)$ as the discounted state distribution, though it does not sum up to one without normalization, which can be achieved by multiplying by $(1-\gamma)$.
 Note that $\mu_\pi$ is equivalent to the discounted state visit count $d^\pi$ introduced by Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806}.

### 2019-03-20 pge-mattis.tex
Fix a sentence that was going horribly wrong
Probably last changes before submission

@@
 Given a differentiable model of the policy, we want to apply a gradient-ascent technique to reach the local optimum. 
-Gradient ascent is theoretically well researched.
+We mainly use gradient ascent, because it is theoretically well researched.
 The main issue is that the policy gradient with respect to the expected return is not available, thus we need to estimate it.
-For this reason, the biggest weakness of policy gradient algorithms is sample efficiency, as they tend to require on-policy data for the gradient estimate.
-Thus, most research is focused on finding algorithms with improved sample efficiency.
+As they also tend to require on-policy data for the gradient estimate, the biggest weakness of policy gradient algorithms is sample efficiency.
+For this reason, most research is focused on finding algorithms with improved sample efficiency.
 This paper provides a deep introduction to policy gradient that shows the development of policy gradient approaches, and should enable the reader to follow current research in the area.
@@
 Using a parameterized policy to decide the next action, they can easily incorporate prior domain knowledge, but require a lot of configuration to produce an effective agent for a specific environment.
-Though they frequently require on-policy training and often converge slower. 
-In this paper, we focus on the computation of policy improvements, though we introduce the most common policy classes shortly.
-In this case, improvement means a step in the parameter space such that the policy under the new parameters will on average perform better than the old policy, i.e. improve its expected return. 
-Policy gradient estimation is the term we use to describe the process of computing the next step in parameter space. 
-The goal is to estimate the gradient of the policy with respect to the expected return. 
+Also, they frequently require on-policy training and a lot of samples to find a good policy. 
+In this paper, we focus on approaches to estimate the policy gradient, though we introduce the most common policy classes shortly.
+Policy improvement means a step in the parameter space such that the policy under the new parameters will on average perform better than the old policy, i.e. improve its expected return. 
+Policy gradient estimation is the term we use to describe the process of computing the direction in parameter space for a policy improvement. 
+Essentially, the goal is to estimate the gradient of the policy with respect to the expected return. 
 Since this is the core problem of policy gradient methods, it is also the main topic of this paper.
@@
 Using our insights from section \ref{sec:pge}, we derive the actor-critic framework in section \ref{sec:ac}, which harnesses value-function estimation for improved gradient updates. 
-Then, in section \ref{sec:natural}, we introduce some gradient-ascent methods that build on the approaches given in \ref{sec:pge}, refining the gradient estimation by Fisher's information matrix to get the natural gradient.
-Finally, in section \ref{sec:outro}, we give a conclusion of the reviewed approaches, and discuss potential areas of improvement.
+Then, in section \ref{sec:natural}, we introduce some gradient-ascent methods that build on the approaches given in sections \ref{sec:pge} and \ref{sec:ac}, refining the gradient estimation by Fisher's information matrix to get the natural gradient.
+Finally, in section \ref{sec:outro}, we summarize the contents presented this paper, and give a short conclusion.
 
@@
 The total reward in the horizon $T$ is $\sum_{t=0}^{T} r_{t}$. 
-We introduce a discount factor $\gamma \in [0,1)$.
+We additionally introduce a discount factor $\gamma \in [0,1)$.
 Intuitively, this reflects the idea that the relevance of later actions declines, and ensures that the return is finite, even for the infinite horizon $T \to \infty$. 
@@
 \end{equation} 
-This can lend a good estimate of the gradient given a small $\delta$, and is generally called the symmetric derivative. 
+This can lend a good estimate of the gradient given a small $\delta \theta$, and is generally called the symmetric derivative. 
 However, finite-difference gradients suffer from the curse of dimensionality, and can require very small $\delta\theta$.
@@
 \paragraph{Value functions.} 
-Given we know the actual value, i.e. the expected return we will get starting from state $s_t$, this function can be used to evaluate the performance of our policy, and can be written as
+Given we know the actual value of a state, i.e. the expected return we will get starting from state $s_t$, this function can be used to evaluate the performance of our policy, and can be written as
 \begin{equation}
@@
 thus $f_w^\pi(s,a)$ being a linear function parameterized by the vector $w$.
-Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806} call this estimator compatible.
+Sutton et al. \cite{Sutton:1999:PGM:3009657.3009806} call this a compatible function approximator.
 This guarantees that the function estimator does not cause divergence, and really enables recent research in reinforcement learning for continuous control problems, e.g., in humanoid robotics.
@@
 However, Peters et al. \cite{4863} find that this approximation is highly reliant on the distribution of the training data. 
-This comes from the realization, that we use only a subspace of the actual action value function in $\hat{V}^\pi$, which is only a state value function. 
+This comes from the realization, that we use only a subspace of the actual action-value function in $\hat{V}^\pi$, which is only a state value function. 
 One can compare this to approximating a parabola by a line, whereby the approximation changes wildly depending on which part of the parabola is in the training data. 
-An approach to solve this bootstrapping problem is to rewrite the Bellman Equation using \eqref{eqn:adv} and \eqref{eqn:q} with $\hat{A}_\pi(s,a) = f_w^\pi(s,a)$, $\hat{V}_\pi(s) = \phi(s)^T v$, and a zero-mean error term $\epsilon(s_t,a_t,s_{t+1})$, we get
+An approach to solve this bootstrapping problem is to rewrite the Bellman Equation using \eqref{eqn:adv} and \eqref{eqn:q}.
+With $\hat{A}_\pi(s,a) = f_w^\pi(s,a)$, $\hat{V}_\pi(s) = \phi(s)^T v$, and a zero-mean error term $\epsilon(s_t,a_t,s_{t+1})$, we get
 \begin{align}
@@
 When using normal gradient steps, we find that steps can become very small when a plateau is reached. 
-This can will drastically slow down the learning process, and in the worst case cause some algorithms to prematurely. 
+This can drastically slow down the learning process, and in the worst case cause algorithms to terminate prematurely. 
 However, we can use some additional information to refine the gradient. 

### 2019-08-07 paper.tex
Convert to IEEE article layout

--- 2019-03-20 pge-mattis.tex
@@
-\begin{filecontents*}{example.eps}
-%!PS-Adobe-3.0 EPSF-3.0
-%%BoundingBox: 19 19 221 221
-%%CreationDate: Mon Sep 29 1997
-%%Creator: programmed by hand (JK)
-%%EndComments
-gsave
-newpath
-  20 20 moveto
-  20 220 lineto
-  220 220 lineto
-  220 20 lineto
-closepath
-2 setlinewidth
-gsave
-  .4 setgray fill
-grestore
-stroke
-grestore
-\end{filecontents*}
 \RequirePackage{fix-cm}
-\documentclass[final]{svjour3}
-\smartqed
+\documentclass[final]{IEEEtran}
 \usepackage[T1]{fontenc}
@@
 \usetikzlibrary{shapes,arrows,positioning,calc}
+
 \begin{document}
 
-\title{Policy Gradient Algorithms}
-\subtitle{An Introduction}
+\title{On Policy Gradients}
 
 \author {Mattis Manfred K{\"a}mmerer}
-
-\institute{TU Darmstadt}
-
 \date{Received: date / Accepted: date}
@@
 This paper provides a deep introduction to policy gradient that shows the development of policy gradient approaches, and should enable the reader to follow current research in the area.
-
-\keywords{reinforcement learning \and policy gradients \and actor-critic \and natural gradient }
 \end{abstract}
@@
 
-\paragraph{Definitions.} 
+\paragraph{Definitions} 
 We define states $s \in \mathbb{S}$, actions $a \in \mathbb{A}$, and rewards $r \in \mathbb{R}$. 
@@
 
-\paragraph{Policy gradient.} 
+\paragraph{Policy gradient} 
 Our goal with respect to episodes is to maximize the expectation of the total reward, also called expected return. 
@@
 
-\paragraph{Finite-difference gradients.} 
+\paragraph{Finite-difference gradients} 
 A simple approach for gradient estimation is to choose a small $\delta\theta$, and evaluate the new policy given the slightly changed parameters as in 
@@
 
-\paragraph{Value functions.} 
+\paragraph{Value functions} 
 Given we know the actual value of a state, i.e. the expected return we will get starting from state $s_t$, this function can be used to evaluate the performance of our policy, and can be written as
@@
 
-\paragraph{Likelihood-ratio gradients.} 
+\paragraph{Likelihood-ratio gradients} 
 For this derivation, we will change the perspective a bit, requiring some additional definitions. 
@@
 
-\paragraph{Episode-based updates.} 
+\paragraph{Episode-based updates} 
 A very general optimization approach to this optimization problem are episodic algorithms. 
@@
 
-\paragraph{Step-based updates.}
+\paragraph{Step-based updates}
 The first class of algorithms developed to update a policy directly using a critic are called REINFORCE \cite{Williams92simplestatistical}.
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{nat-grad-adv}
+  \includegraphics[width=0.5\textwidth]{nat-grad-adv}
   \caption{An experiment showing where the natural gradient has a great advantage \cite{Peters_IICHR_2003}. }\label{fig:nat-grad-adv}
@@
 \begin{equation}
-  \nabla_\theta J(\theta) = \int_\mathbb{S} {d^\pi(s) \int_\mathbb{A} \pi(a|s) \nabla_\theta \log \pi(a|s) {\nabla_\theta \log \pi(a|s)}^T dads}\ w \equiv F_\theta w.
+  \nabla_\theta J(\theta) = F_\theta w.
   \label{eqn:J-equals-F}

### 2019-08-07 paper.tex
Use present tense

@@
 When we take out this common factor, $Q_\pi(s,a)$ remains.
-Then, we used $\nabla_\theta \pi(a|s) = \pi(a|s)\nabla_\theta\log\pi(a|s)$, obtained from the likelihood ratio $\nabla_\theta \log p(x|\theta) = \frac{\nabla_\theta p(x|\theta)}{p(x|\theta)}$. 
+Then, we use $\nabla_\theta \pi(a|s) = \pi(a|s)\nabla_\theta\log\pi(a|s)$, obtained from the likelihood ratio $\nabla_\theta \log p(x|\theta) = \frac{\nabla_\theta p(x|\theta)}{p(x|\theta)}$. 
 This gives us the likelihood-ratio gradient

### 2019-08-07 paper.tex
Fix typo
Improve formatting

@@
 
-\paragraph{Definitions} 
 We define states $s \in \mathbb{S}$, actions $a \in \mathbb{A}$, and rewards $r \in \mathbb{R}$. 
@@
 An approach to solve this bootstrapping problem is to rewrite the Bellman Equation using \eqref{eqn:adv} and \eqref{eqn:q}.
-With $\hat{A}_\pi(s,a) = f_w^\pi(s,a)$, $\hat{V}_\pi(s) = \phi(s)^T v$, and a zero-mean error term $\epsilon(s_t,a_t,s_{t+1})$, we get
+With $\hat{A}_\pi(s,a) = f_w^\pi(s,a)$, $\hat{V}_\pi(s) = \phi(s)^T v$, a zero-mean error term $\epsilon \equiv \epsilon(s_t,a_t,s_{t+1})$, we get
 \begin{align}
-  \hat{Q}_\pi(s,a) = \hat{A}_\pi(s,a) + \hat{V}_\pi(s) &= r(s,a) + \gamma \int_\mathbb{S} p(s'|s,a)\hat{V}_\pi(s')ds', \\
-  \nabla_\theta \log \pi(a_t|s_t)^T w + \phi(s_t)^T v &= r(s_t,a_t) + \gamma \phi(s_{t+1})^T v + \epsilon(s_t,a_t,s_{t+1}).
+  &\hat{A}_\pi(s,a) + \hat{V}_\pi(s) = r(s,a) + \gamma \int_\mathbb{S} p(s'|s,a)\hat{V}_\pi(s')ds', \\
+  &\nabla_\theta \log \pi(a_t|s_t)^T w + \phi(s_t)^T v = r(s_t,a_t) + \gamma \phi(s_{t+1})^T v + \epsilon ,
 \end{align}
-which gives involves only linear equations to solve \cite{4863}.
+which involves only linear equations to solve \cite{4863}.
 
@@
 \begin{figure}
-  \includegraphics[width=0.5\textwidth]{nat-grad-adv}
+  \includegraphics[width=0.485\textwidth]{nat-grad-adv}
   \caption{An experiment showing where the natural gradient has a great advantage \cite{Peters_IICHR_2003}. }\label{fig:nat-grad-adv}

### 2019-08-15 paper.tex
Update some details

@@
 
-\author {Mattis Manfred K{\"a}mmerer}
-\date{Received: date / Accepted: date}
+\author {Mattis Manfred K{\"a}mmerer \\ \normalsize{Technische Universit{\"a}t Darmstadt} \\ \normalsize{mattis.kaemmerer@gmx.de}}
 
@@
 The goal of policy gradient approaches is to find a policy in a given class of policies which maximizes the expected return.
-Given a differentiable model of the policy, we want to apply a gradient-ascent technique to reach the local optimum. 
+Given a differentiable model of the policy, we want to apply a gradient-ascent technique to reach a local optimum. 
 We mainly use gradient ascent, because it is theoretically well researched.
 The main issue is that the policy gradient with respect to the expected return is not available, thus we need to estimate it.
-As they also tend to require on-policy data for the gradient estimate, the biggest weakness of policy gradient algorithms is sample efficiency.
+As policy gradient algorithms also tend to require on-policy data for the gradient estimate, their biggest weakness is sample efficiency.
 For this reason, most research is focused on finding algorithms with improved sample efficiency.
-This paper provides a deep introduction to policy gradient that shows the development of policy gradient approaches, and should enable the reader to follow current research in the area.
+This paper provides a formal introduction to policy gradient that shows the development of policy gradient approaches, and should enable the reader to follow current research on the topic.
+
 \end{abstract}
@@
 Then, in section \ref{sec:natural}, we introduce some gradient-ascent methods that build on the approaches given in sections \ref{sec:pge} and \ref{sec:ac}, refining the gradient estimation by Fisher's information matrix to get the natural gradient.
-Finally, in section \ref{sec:outro}, we summarize the contents presented this paper, and give a short conclusion.
+Finally, in section \ref{sec:outro}, we summarize the contents presented in this paper, and give a short conclusion.
 
@@
 \end{align}
-Since $\mu_0^T$ does not depend on $\theta$, it is obvious that 
+Since $\mu_0^T$ does not depend on $\theta$, 
 \begin{align*}
@@
 \end{equation}
-intuitively meaning we should to increase the probability of actions that lead to higher Q-values.
+intuitively meaning we should increase the probability of actions that lead to higher Q-values.
 This formulation enables us to calculate the gradient $\nabla_\theta J(\theta)$, while directly taking advantage of the MDP structure in the form of $Q_\pi(s,a)$.
@@
 However, Peters et al. \cite{4863} find that this approximation is highly reliant on the distribution of the training data. 
-This comes from the realization, that we use only a subspace of the actual action-value function in $\hat{V}^\pi$, which is only a state value function. 
+This comes from the realization, that we use only a subspace of the true action-value function in $\hat{V}^\pi$, which is only a state value function. 
 One can compare this to approximating a parabola by a line, whereby the approximation changes wildly depending on which part of the parabola is in the training data. 
@@
 of the natural gradient. 
-The Fisher information matrix represents the certainty we have on our estimate of the gradient and is defined as the covariance of the log likelihood function of a trajectory $\tau_{\pi}^T$, which as \cite{4863} show can be written as
+The Fisher information matrix represents the certainty we have on our estimate of the gradient and is defined as the covariance of the log likelihood function of a trajectory $\tau_{\pi}^T$, which as Peters et al.
+\cite{4863} show can be written as
 \begin{equation}
@@
 In this paper, we have introduced policy gradient methods as a class of reinforcement learning algorithms. 
-We show why policy gradient methods are effective in these environments, and we give some intuitions about the concept. 
+We show why policy gradient methods are effective in these environments, and we give some intuitions for the concept. 
 Further, we show the core elements of policy gradient methods, discuss some intricacies the estimation of the policy gradient brings, and follow the research development in the attempts of improving the efficiency and stability of policy gradients.
